---
title: "Climatic_data"
author: "Thomas Francisco"
date: "2024-04-29"
output:
  html_document:
    number_sections: true #titles
    toc: true #table of content
    toc_float: true # enable the toc to be on the side of the text, always visible
    collapsed: True #control if the toc label will only display top level titles
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)

#rm(list = ls()) 
library(dplyr) 
library(tidyverse)
library(ggplot2)
library(corrplot)  
library(dismo)
library(raster)
library(vegan)
library(plyr)
library(writexl)
library(rasterize)
```

# Introduction

This script processes climatic and environmental data from the raw output of *ClimateDT* (or any program that provides data in raster TIFF format) into a dataframe containing standardized data for the selected variables for past, present, and future climatic conditions. Climatic data at high resolution (30 arc-seconds) were extracted from the Climate Downscaling Tool (ClimateDT; Marchi et al., 2024). ClimateDT is a dynamic downscaling web tool for elevation-adjusted historical and future climatic data. The period from 1901 to 1950 was defined as the reference period, from which the populations are currently locally adapted. We selected the period from 2041 to 2060 as the future timeframe and used the mean values from five global climate models (GCMs) under the socio-economic pathway 3-7.0 for our climatic predictions.

We performed several selection steps to identify the final set of climatic predictors used in downstream analyses. First, the climatic selection was conducted on a pre-selected set of climatic variables identified as relevant to the species' ecology. For past climatic data, we performed a forward selection of variables to identify those most closely related to genomic variation across populations, using the ordiR2step function from the vegan R package (Oksanen et al., 2001). Next, we calculated the correlation between each variable in the initial set of climatic predictors and removed variables that were over-collinear (correlation coefficient > 0.75), while retaining, when possible, the most informative climatic variables identified during the forward selection. Finally, we calculated a variance inflation factor (VIF) for each predictor and removed additional climatic variables to adjust the VIF values below 10.

After the selection steps, we standardized the climatic data and output the retained climatic data for past, present, and future scenarios in several formats.

```{r meta data}
meta_data_pop <- read.csv("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Populations/taxus_sample_29pop.csv",h=T,sep=";",dec=",")
meta_data_vcf <- read.csv("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/samples/samples_taxus_baccata_adapcon_gentree.csv",h=T,sep=";",dec=",")
```

# Pre-selection of climatic variables


The first step consisted of reviewing the literature on the targeted species to define a set of climatic variables that could potentially drive (induce selective pressure on) adaptive genetic variation across the species range. The goal is to retain only biologically informative climatic variables. This approach reduces the number of climatic variables tested and allows us to refine our hypotheses before starting the analysis. The objective is not to limit our analyses to only 2 or 3 variables, nor to retain 40; this pre-selection could include 10-15 variables, depending on the species and the richness of the relevant literature.

For Taxus baccata, the initial pre-selection includes: **Bio 1, 2, 3, 4, 5, 6, 7, 9, 10, 12, 13, 14, 15, 16, 17, 18, MGSP, AHM, SHM**.
  
# Climatic data

## Data

The climatic data are obtained from *ClimateDT* in two different formats: point layers and raster. We extracted data for both past and future conditions. One major difference between the past and future data lies in their structure:

Past data are available as one value per year for each variable (for each selected population/pixel). For example, for the period 1901-1950, we have 50 values for each variable for each population.
Future data are less detailed, providing only the mean values for the selected period. For instance, for the period 2041-2070, we have just one value for each variable for each population.
One of the challenges is that all bioclimatic variables (bio) or other related climatic variables are constructed using monthly values of minimum temperature (tmin), maximum temperature (tmax), and precipitation. We cannot simply calculate the mean of the 50 values of Bio 1 from 1901 to 1950 to obtain the average for that period. Instead, we need to average the tmin, tmax, and precipitation for each month to obtain one tmin value for January for the period 1901-1950, one for February, and so on.

We can then use the *biovars* function from the **Dismo** R package to calculate the 19 bioclimatic variables for the period of interest (e.g., 1901-1950). To calculate other climatic variables such as AHM for the same specific period, we can follow the *ClimateDT* procedures using the new Bio1, Bio10, and Bio12 values obtained from the *biovars* function.

For other variables not directly linked to tmin, tmax, precipitation, or bioclimatic variables, such as GDD5, obtaining the average value for a specific period is not straightforward, as the data come from Maurizio Marchi (IBBR-CNR).

Additionally, only one model is available in ClimateDT for future data. It is important to note that climatic models can predict different values depending on the method used to estimate future climate. Therefore, we could consider using the mean future climatic values from several climatic models to avoid selecting an extreme model across the range of predicted values. To utilize multiple climatic models, we can only access raster data.

**We have already checked this period in the script Processing_climatic_data_T_adapcon_Gentree_bis and tested for differences between point layers and raster data.**

Based on these previous scripts, **we will use the 1901-1950 period and raster data because the differences between raster values and point layers are significant for temperature-related climatic variables. Indeed, for future data, we need to use raster, as explained earlier, so it is preferable to also use raster data for past climatic data to avoid differences in climate linked to the method of extraction.**

Load past climatic data 
```{r load past raster data}
#create an R object where all the climatic rasters are.
ras.bio <- stack(list.files("C:/Users/tfrancisco/Documents/Thesis/Data/Climatic/1901-1950_raster/", pattern = ".tif$", full.names = T)) 
#names of the raster (because they are not in the classic order Bio 1, 2 etc they are order by number so bio 1, 10, 11 etc)
names(ras.bio) <- unlist(strsplit(unlist(lapply(strsplit(list.files("C:/Users/tfrancisco/Documents/Thesis/Data/Climatic/1901-1950_raster/", pattern = ".tif$", full.names = T), split = "C:/Users/tfrancisco/Documents/Thesis/Data/Climatic/1901-1950_raster/"), function(x) x[2])), split = ".tif"))

#we extracted the climatic values for each populations from the raster based on their coordinates
coords<- data.frame(apply(meta_data_pop[,c(5:4)], 2, as.numeric))#we need to have longitude then latitude
past_climatic_data_raster <- data.frame(meta_data_pop$Population,raster::extract(ras.bio, coords))#important to add raster:: because this function is also in tidyr and will not do the same things
colnames(past_climatic_data_raster)<- c("Population",names(ras.bio))

#add country info 
past_climatic_data_f <- merge(meta_data_pop[,c(1,2)],past_climatic_data_raster, "Population")
```


```{r save past raster for GO RDA, include=FALSE}
#We need to save the raster for RDA GO
past_climatic_data_raster <- ras.bio
save(past_climatic_data_raster,file="C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Genomic_offset/RDA/past_climatic_data_raster.Rdata")
```

## Climatic variations among populations
      
```{r violin plot in a loop}
#do the violin plot in a loop for each climatic variables

#input data for the loop
data <- past_climatic_data_f
regions <-past_climatic_data_f$Country
bioclim_vars <- c("bio1", "bio2", "bio3", "bio4","bio5","bio6","bio7","bio9", "bio10","bio12","bio13","bio14","bio15","bio16", "bio17", "bio18","AHM", "MGSP", "SHM", "GDD5")
unit_var <- c("°c","°c","index","°c","°c","°c","°c","°c","°c","mm","mm","mm","index","mm","mm","mm","°c/mm","mm","°c/mm","°c*days" )

#results: the loop
for (x  in 1:length(bioclim_vars)) {
  var <- bioclim_vars[x]
   unit <- unit_var[x]
  data_subset <- data.frame(data[, paste0(var)]) #for each variable in bioclim_vars, we extract the two periods and created a data.frame with only these two variables. 
  colnames(data_subset) <- c("var1")
  # Create scatterplot to compare the bioclimatic variable between the two periods
  plot <- ggplot(data_subset, aes(x =var,y= var1)) +#var is the values and var1 is just the name of the variable 
  geom_violin(trim = FALSE) +
  geom_point(aes(color = regions), shape = 16, position = position_jitter(seed = 1, width = 0.2)) +
  labs(colour="Countries",x="variable",y=paste0("values ","(",unit,")"),title=paste0("Violin Plot of ",var," differences across populations"))+
  theme(plot.title = element_text(hjust = 0.5))
  
  # Add any additional customization or saving the plot if needed
png(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/climate/past_climatic_data/violin_plot/violin_plot_",var,".png"));print(plot)
;dev.off()
}
```

Example of violin plot: 

```{r example of violin plot for the html, echo=FALSE}
data <- past_climatic_data_f
regions <-past_climatic_data_f$Country
bioclim_vars <- c("bio1", "bio4","bio12","bio15")
unit_var <- c("°c","°c","mm","mm" )

#results: the loop
for (x  in 1:length(bioclim_vars)) {
  var <- bioclim_vars[x]
   unit <- unit_var[x]
  data_subset <- data.frame(data[, paste0(var)]) #for each variable in bioclim_vars, we extract the two periods and created a data.frame with only these two variables. 
  colnames(data_subset) <- c("var1")
  # Create scatterplot to compare the bioclimatic variable between the two periods
  plot <- ggplot(data_subset, aes(x =var,y= var1)) +#var is the values and var1 is just the name of the variable 
  geom_violin(trim = FALSE) +
  geom_point(aes(color = regions), shape = 16, position = position_jitter(seed = 1, width = 0.2)) +
  labs(colour="Countries",x="variable",y=paste0("values ","(",unit,")"),title=paste0("Violin Plot of ",var," differences across populations"))+
  theme(plot.title = element_text(hjust = 0.5))

# Print the plot
  print(plot)
}
```

Interpretation: We observed differences in climatic values across populations. This could suggest that local adaptation is present, as there appear to be varying climatic selective pressures affecting the different populations.  
Conclusion: These findings imply that local adaptation may be occurring, given the distinct climatic selective pressures across populations.   
  
# Selection of climatic data
  
The selection of climatic variables was considered through five steps:  
  
##Difference between past and future values

We will not remove variables that do not change in the future because if certain variables remain stable, it indicates that climate change may not significantly impact these species or populations. This is particularly relevant for populations that are locally adapted to conditions that will not change.  

## Imprecision along studied area
      
The accuracy of climatic variables is not homogeneous throughout the studied area or among the climatic variables themselves. If layers of uncertainty are available, we could assess whether some variables exhibit high inaccuracy across the study area and remove them if necessary. However, this is not feasible for now.
 
## Explained variance

In this step, we aim to identify which variables explain the most genetic variation. To achieve this, we conducted a Redundancy Analysis (RDA; see the explanation of the method in the script RDA_candidate_detection). Specifically, we used the *ordiR2step* function, starting with a null model that includes no explanatory variables and progressing to a full model that incorporates all potential explanatory variables. The stopping criterion was the significance of a variable in increasing the R² of the model, using a threshold of p-value < 0.05 based on 1000 permutations. From the p-values obtained, we identified the most important variables that explain genomic variation. Twenty runs were conducted to evaluate the consistency of the detection of these significant variables.

```{r genomic data}
#genomic data
load("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/genetic_new/data_allelic_frequencies_29pop_adapcon_gentree_475_8616.Rdata")
genomic_matrix <- data_allelic_frequencies_29pop_adapcon_gentree_475_8616
```

We also need to scale the climatic var: 
```{r scale}
past_climatic_data_scale <- past_climatic_data_f[,-c(1,2)]%>%
  apply(2,as.numeric) %>% as.data.frame %>% 
mutate(across(where(is.numeric), scale))
```

We can perform the RDA:
```{r RDA, eval=F, echo=T}
#null model
RDA_null <- rda(genomic_matrix~1,past_climatic_data_scale )

#model with all the variables
RDA_full <- rda(
formula=genomic_matrix~bio1+bio2+bio3+bio4+bio5+bio6+bio7+bio9+bio10+bio12+bio13+bio14+bio15+bio16+bio17+bio18+AHM+MGSP+SHM+GDD5,data = past_climatic_data_scale,scale=F)

#stepwise by selecting only variables that increase the adjusted R^2 of the model and that have a pvalue lower than 0.05
Variable_selection_pv0.05 <- ordiR2step(RDA_null, RDA_full, Pin = 0.05, R2permutations = 1000, R2scope = T)

#Variable_selection_pv0.05$anova

#we want, following Archambeau et al 2024, investigate if the permutation stepwise identified the same variables of interest depending on the run:
#we run 20 independent stepwise models.
nbmodels <- 20

# Stepwise selection with ordiR2step function
rep20_variable_selec_pv0.05 <- lapply(1:nbmodels, function(x) {
  mod <- ordiR2step(RDA_null, RDA_full, Pin = 0.05, R2permutations = 1000, R2scope = T)
  return(names(mod$CCA$envcentre))}) %>% 
  setNames(paste0("model",1:nbmodels)) %>% 
  ldply(function(x) data.frame(variables=x),.id="models") %>%  
  dplyr::summarise(count(variables)) %>% 
  setNames(c("variable","count"))
```

```{r save and load rep ordi2steps, include=FALSE}
#save(rep20_variable_selec_pv0.05, file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/climate/selection_var/rep20_variable_selec_pv0.05.Rdata")

load("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/climate/selection_var/rep20_variable_selec_pv0.05.Rdata")
```

```{r see rep20_variable_selec_pv0.05}
rep20_variable_selec_pv0.05
```

Interpretation: Bio 2, 4, 7 and 9 appear to be important climatic variables. They were consistently identified in 20 out of 20 runs of the ordiR2step analysis, each contributing to an increase in the R² of the models with p-values < 0.05.
Conclusion: According to the stepwise analysis, **bio 2, 4, 7 and 9** seems to be of interest to explain the genetic variation. 
      

## Multicollinerity between variables
      
Here, we aim to select variables with low collinearity. To achieve this, we retained only those variables that are not highly correlated (with an absolute correlation coefficient < 0.75) from the pre-selected set of climatic variables.

```{r multicollinearity between variables}
#function to do it
#matrix of correlation
correlation_function <-function(data,threshold){
    data_correlation <- subset(data,select= -c(Population,Country)) 
      rownames(data_correlation) <- data$Populations
      correlation <- cor(data_correlation)
  correlation[abs(correlation) <= threshold] <- 0
corr_plot <- corrplot(correlation, method = "number", addrect = 2, col = c("red", "white", "red"), type = "lower", tl.col = "black", tl.cex = 0.6, number.cex = 0.6)
#save
png("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/climate/selection_var/corrplot_multicollinearity.png");corr_plot <- corrplot(correlation, method = "number", addrect = 2, col = c("red", "white", "red"), type = "lower", tl.col = "black", tl.cex = 0.6, number.cex = 0.6);dev.off()

#plot the corrplot
corr_plot
}

#right order
past_climatic_data_order <- past_climatic_data_f[,c(1,2,4,13:19,5:12,3,20,21,22)]

#correlation past/present
data_present <- past_climatic_data_order #we use the dataset without the variables we discarded previously
threshold <- 0.75
correlation_past <-correlation_function(data_present,threshold)
```

Conclusion: Based on the correlation between the variables and the prior information on the most important variables explaining the genetic variation, we can select **BIO 1, 2, 3, 4, 5, 9, 12, 15**

## Variance inflation factor (VIF) 
      
We aim to test the inflation factor among the retained variables to ensure that we do not include overcorrelated variables.
```{r VIF}
#subset step 4
RDA_subset_step4 <- rda(formula=genomic_matrix~bio1+bio2+bio3+bio4+bio5+bio9+bio12+bio15,data = past_climatic_data_scale,scale=F)
#anova.cca(RDA_subset_step4)
#RsquareAdj(RDA_subset_step4)
##vif
vif.cca(RDA_subset_step4)
```

We observe that the VIF for some variables in the subset from step 4 is excessively high (over 160 for BIO 2). Based on this observation, along with the understanding that certain variables are important for explaining genetic variation, we adjusted the subset to ensure that the VIF values are below 10.

```{r final set}
#adjusted subset based on VIF
RDA_adjusted_subset <- rda(formula=genomic_matrix~bio1+bio2+bio4+bio9+bio12+bio15,data = past_climatic_data_scale,scale=F)
#anova.cca(RDA_adjusted_subset)
#RsquareAdj(RDA_adjusted_subset)
##vif
vif.cca(RDA_adjusted_subset)

```

Interpretation: By removing BIO 5 and BIO 3, we successfully lowered the VIF for all retained variables to less than 3 

Conclusion: we retained the subset: **Bio 1, 2, 4, 9, 12, 15** :
    - Bio 1: Mean annual Temperature (°C)
    - Bio 2: Mean Diurnal Range (Mean of monthly (max temp - min temp)) (°C)
    - Bio 4: Temperature Seasonality (standard deviation x100) (°C)
    - Bio 9: Mean Temperature of Driest Quarter (°C)
    - Bio 12: Mean annual Precipitation (mm)
    - Bio 15: Precipitation Seasonality (mm)
    
    
# Retained climatic data other periods

Based on the climatic selection, we can keep for present and future periods, these variables: **Bio 1, 2, 4, 9, 12, 15**

## Present

```{r load present raster data}
#create an R object where all the climatic rasters are.
ras.bio <- stack(list.files("C:/Users/tfrancisco/Documents/Thesis/Data/Climatic/1991-2020_raster_climateDT/", pattern = ".tif$", full.names = T)) 
#names of the raster (because they are not in the classic order Bio 1, 2 etc they are order by number so bio 1, 10, 11 etc)
names(ras.bio) <- unlist(strsplit(unlist(lapply(strsplit(list.files("C:/Users/tfrancisco/Documents/Thesis/Data/Climatic/1991-2020_raster_climateDT/", pattern = ".tif$", full.names = T), split = "C:/Users/tfrancisco/Documents/Thesis/Data/Climatic/1991-2020_raster_climateDT/"), function(x) x[2])), split = ".tif"))

#we extracted the climatic values for each populations from the raster based on their coordinates
coords<- data.frame(apply(meta_data_pop[,c(5:4)], 2, as.numeric))#we need to have longitude then latitude
present_climatic_data_Adapon_pop <- data.frame(meta_data_pop$Population,raster::extract(ras.bio, coords))#important to add raster:: because this function is also in tidyr and will not do the same things
colnames(present_climatic_data_Adapon_pop)<- c("Population",names(ras.bio))

#add country info 
present_climatic_data_Adapon_pop_f <- merge(meta_data_pop[,c(1,2)],present_climatic_data_Adapon_pop, "Population")

#order
Present_new_6_Climatic_data_scale <-present_climatic_data_Adapon_pop_f[,c(1,2,3,6:8,4,5)]

#name bio 
colnames(Present_new_6_Climatic_data_scale) <- c("Population","Country","Annual_Tc","Diurnal_range_Tc","Tc_Seasonality","Tc_driest_quarter","Annual_P","P_Seasonality")

```


For the gradient forest genomic offset, we will need to extract the climatic conditions for the clonal bank for each population, resulting in one CSV file where the "future" conditions correspond to those at the clonal bank site.

## Future climatic data 

For future data, multiple climatic models are available. Choosing one of them without prior information could lead to imprecision, as the differences in predictions can sometimes be significant. To avoid bias from selecting only one climatic model that could be extreme, we selected five models for the future data.

We want to compare these models to assess the differences in predictions. For future climatic data, we chose the 2041-2070 period under the SSP 3.7.0 scenario. Since we did not have any prior information to guide the choice of models, we selected five of the most commonly used climatic models:
    - GFDL-ESM4  
    - IPSL-CM6A-LR  
    - MPI-ESM1-2-HR 
    - MPI-ESM2-0  
    - UKESM1-0-LL  

The first step is to load the future climatic data for each climatic model
```{r load future climatic data}
#list of the model for the loop
list_models<- c("GFDL_ESM4","IPSL_CM6A_LR","MPI_ESM1_2_HR","MPI_ESM2_0","UKESM1_0_LL") 

for(x in 1:length(list_models)){
  
  model <- list_models[x]
  #create an R object where all the climatic rasters are
ras.bio <- stack(list.files(paste0("C:/Users/tfrancisco/Documents/Thesis/Data/Climatic/Future_climate/2041_2070_",model,"_ssp370/"), pattern = ".tif$", full.names = T)) 
#names of the raster (because they are not in the classic order Bio 1, 2 etc they are order by number so bio 1, 10, 11 etc)
names(ras.bio) <- unlist(strsplit(unlist(lapply(strsplit(list.files(paste0("C:/Users/tfrancisco/Documents/Thesis/Data/Climatic/Future_climate/2041_2070_",model,"_ssp370/"), pattern = ".tif$", full.names = T), split = paste0("C:/Users/tfrancisco/Documents/Thesis/Data/Climatic/Future_climate/2041_2070_",model,"_ssp370/")), function(x) x[2])), split = ".tif"))

future_climatic_data <- data.frame(meta_data_pop$Population,raster::extract(ras.bio, coords))
colnames(future_climatic_data)=c("Population",names(ras.bio))
new_colnames<- paste(colnames(future_climatic_data[,-1]),model, sep = "_")
colnames(future_climatic_data)[-1] <- new_colnames

#add country info 
future_climatic_data_f <- merge(meta_data_pop[,c(1,2)],future_climatic_data, "Population")

assign(paste0("Future_data_",model), future_climatic_data_f)
}
```


The next step is to group by climatic variables the data across climatic models
```{r creata 1 df per bio}
#list of the climatic var
list_climatic_var <- c("bio1","bio12","bio15","bio2","bio4","bio9")

for(x in 1:length(list_climatic_var)){
  climatic_var <- list_climatic_var[x]
  
  pos_clim <- x+2
  
  #climatic_values_ff <- Future_data_GFDL_ESM4[,x]
  
  data_set_clim <- data.frame(Population=Future_data_GFDL_ESM4$Population,Country=Future_data_GFDL_ESM4$Country,Future_data_GFDL_ESM4[,pos_clim],Future_data_IPSL_CM6A_LR[,pos_clim],Future_data_MPI_ESM1_2_HR[,pos_clim],Future_data_MPI_ESM2_0[,pos_clim],Future_data_UKESM1_0_LL[,pos_clim])
  
  colnames(data_set_clim)<- c("Population","Country",paste0(climatic_var,"_GFDL_ESM4"),paste0(climatic_var,"_IPSL_CM6A_LR"),paste0(climatic_var,"_MPI_ESM1_2_HR"),paste0(climatic_var,"_MPI_ESM2_0"),paste0(climatic_var,"_UKESM1_0_LL"))
  
  assign(paste0(climatic_var,"_future_climatic_data"),data_set_clim)
}
```


Then, we can generate violin plots of the climatic values for each population using each climatic model. We will generate one graph for each bioclimatic variable of interest. 
```{r plot the variation across models, message=FALSE, warning=FALSE}

#input data for the loop
unit_var <- c("°c","mm","Coefficient of Variation","°c","sd°c x100","°c")
list_dataset <- c("bio1_future_climatic_data","bio12_future_climatic_data","bio15_future_climatic_data","bio2_future_climatic_data","bio4_future_climatic_data","bio9_future_climatic_data")

#results: the loop
for (x  in 1:length(list_dataset)) {
  
  dataset <- get(list_dataset[x])
   unit <- unit_var[x]
var_name <- list_climatic_var[x]
  
data_violin <- gather(dataset,key="dataset",value="Value",-Population,-Country)

#add the name of the model 
data_violin_f <- data_violin %>%
  mutate(models = gsub("^[^_]+_", "", dataset))


  # Create scatterplot to compare the bioclimatic variable between the two periods
  plot <- ggplot(data_violin_f, aes(x = models,y= Value)) +
  geom_violin(trim = FALSE) +
  geom_point(aes(color = data_violin_f$Country), shape = 16, position = position_jitter(seed = 1, width = 0.2)) +
  labs(colour="Regions",x=var_name,y=paste0("values ","(",unit,")"),title=paste0("Violin Plot of ",var_name," differences across climatic models"))+
  theme(plot.title = element_text(hjust = 0.5)) +
     scale_x_discrete(labels = c("label1" = "name1", "label2" = "name2")) 
  
#save
png(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/climate/future_climatic_data/violin_plot_comp_models/Comparison_",var_name,"between_climatic_models.png"));print(plot)
;dev.off()

# Print the plot
  print(plot)
}
```

We can see that there are some differences between the climatic models, but overall, the values seem quite close. To test if the values are similar, we will calculate a correlation matrix.

```{r correlation values across climatic models, message=FALSE, warning=FALSE}

for(x in 1:length(list_dataset)){
  df <- get(list_dataset[x])
  
  df_num <- df[,-c(1,2)] %>% mutate(across(as.numeric))
  
  #names of the corr matrix for each biovariable
  names_corr <- paste0("correlation_",list_dataset[x])
  #title of corrplot
  title <- (paste0("Correlation ",list_dataset[x]))
  
  
  
  #group for each bio var with the values of Bayes factor only 
corr_bio <- cor(df[, grepl("bio", names(df))]) 

#name the corr_bio with names_corr
assign(names_corr,corr_bio)

#plot corrplot
corr_plot <- corrplot(get(names_corr), method = "number", addrect = 2, col = c("red", "white", "red"), type = "lower", tl.col = "black", tl.cex = 0.6, number.cex = 0.6, title = title,mar=c(0,0,1,0) )

#save
png(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/climate/future_climatic_data/Correlation/correlation_",list_dataset[x],".png"));corr_plot <- corrplot(get(names_corr), method = "number", addrect = 2, col = c("red", "white", "red"), type = "lower", tl.col = "black", tl.cex = 0.6, number.cex = 0.6, title = title,mar=c(0,0,1,0) );dev.off()
}
```

The climatic values are very similar across climatic models, except for **BIO 9**, where the **UKESM1-0-LL model** predicts values that are very inconsistent with those predicted by the other models. This serves as a reminder that, in the absence of prior information suggesting one model over the others, we should be cautious when selecting future climatic values. Here, we will use the mean values of the five climatic models as our future values.

```{r mean values across climatic models}

for (x in 1: length(list_dataset)){
  var <- list_climatic_var[x]
   dataset <- get(list_dataset[x]) 

df <- dataset %>%
  mutate(mean_values=rowMeans(dplyr::select(., 3:7)))

df_f <- df[,c(1,2,8)]
colnames(df_f) <- c("Population","Country",paste0("mean_values_",var))

assign(paste0("df_",var,"_mean"),df_f)
}
```

```{r merge in 1 dataframe}

Future_climatic_data_T_adapcon_gentree<- data.frame(Population=df_bio1_mean[,1],Country=df_bio1_mean[,2],Annual_Tc=df_bio1_mean[,3],Diurnal_range_Tc=df_bio2_mean[,3],Tc_Seasonality=df_bio4_mean[,3],Tc_driest_quarter=df_bio9_mean[,3],Annual_P=df_bio12_mean[,3],P_Seasonality=df_bio15_mean[,3])
```


# Standardization of the retained climatic variables
  
The final step is to create a data frame with the retained climatic variables standardized for further analysis. 
We standardized the present and future data using the past values to enable comparisons of indices based on the climatic data between periods (e.g., Adaptive Index, Genomic Offset, etc.).

## Past data 

```{r normalize the past data, message=FALSE, warning=FALSE}
#dataframe not scaled
Past_clim_data_Europe_pop <- past_climatic_data_order[,c(1:4,6,10,12,15)]
#creation of the scaled matrix
Past_new_6_Climatic_data_scale <-past_climatic_data_order[,c(3,4,6,10,12,15)] %>% 
    scale()


Past_new_6_Climatic_data_scale_df <- data.frame(past_climatic_data_order[,c(1,2)],Past_new_6_Climatic_data_scale)

colnames(Past_new_6_Climatic_data_scale_df) <- c("Population","Country","Annual_Tc","Diurnal_range_Tc","Tc_Seasonality","Tc_driest_quarter","Annual_P","P_Seasonality")


df_to_scale <- past_climatic_data_order[,c(3,4,6,10,12,15)]
colnames(df_to_scale) <- c("Annual_Tc","Diurnal_range_Tc","Tc_Seasonality","Tc_driest_quarter","Annual_P","P_Seasonality")

  scale_env_value_new_cli <- attr(scale(df_to_scale), 'scaled:scale')
center_env_value_new_cli <- attr(scale(df_to_scale), 'scaled:center') 
```

```{r save past climatic data, include=FALSE}
#Save the matrix and the normalized values
#non scaled data
save(Past_clim_data_Europe_pop,file="C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/Past_clim_data_Europe_pop.Rdata" )

#scaled data
write_xlsx(Past_new_6_Climatic_data_scale_df,"C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/Past_new_6_Climatic_data_scale_df.xlsx")
save(Past_new_6_Climatic_data_scale_df,file="C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/Past_new_6_Climatic_data_scale_df.Rdata")

#Save the values used to scaled
save(scale_env_value_new_cli,file="C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/scale_env_value_new_cli.Rdata",force=T)
save(center_env_value_new_cli,file="C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/center_env_value_new_cli.Rdata",force=T)
```


We also need to save the climatic data into a raster: 
```{r data_to_be_used_raster}
Past_new_6_Climatic_data_df <- data.frame(past_climatic_data_order[,c(1,2)],past_climatic_data_order[,c(3,4,6,10,12,15)])

past_clim_data <- data.frame(merge(meta_data_pop[,c(2,4,5)],Past_new_6_Climatic_data_df[,c(1,3,7,8,4:6)],"Population"))
# we need to rename colnames for raster
past_clim_data$Latitude <- as.numeric(past_clim_data$Latitude);past_clim_data$Longitude <- as.numeric(past_clim_data$Longitude)
Past_clim_data_new <- data.frame(past_clim_data[,c(3,2)],past_clim_data[,c(4:9)])

colnames(Past_clim_data_new) <- c('x', 'y',"bio1", "bio12", "bio15","bio2","bio4","bio9")
```


```{r rasterization past data}
# create a raster object with the extent of the data
r_obj <- raster(xmn=min(Past_clim_data_new$x)-0.1, xmx=max(Past_clim_data_new$x)+0.1, ymn=min(Past_clim_data_new$y)-0.1, ymx=max(Past_clim_data_new$y)+0.1, resolution=c(0.01,0.01)) #the resolution is smaller than what we have in meta_data_pop

# use rasterize to create desired raster with the pop and their values
past_climatic_data_raster <- rasterize(x=Past_clim_data_new[, 1:2], # lon-lat data
                    y=r_obj, # raster object
                    field=Past_clim_data_new[, 3:8], # clim var to fill raster with
                    fun=mean) # aggregate function
```

```{r save raster past data, include=FALSE}
save(past_climatic_data_raster,file="C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Genomic_offset/RDA/Past_climatic_data_raster.Rdata",F=T)
```


## Present

```{r present data}
#scale 
load("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/scale_env_value_new_cli.Rdata")
load("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/center_env_value_new_cli.Rdata")

Present_climatic_data_T_adapcon_gentree_scaled <- data.frame(Present_new_6_Climatic_data_scale[,c(1,2)],scale(Present_new_6_Climatic_data_scale[,-c(1,2)], center = center_env_value_new_cli, scale = scale_env_value_new_cli))
```

We need to save the raster for RDA GO
```{r save present raster for GO RDA, include=FALSE}
#raster
present_climatic_data_raster <- ras.bio
save(present_climatic_data_raster,file="C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Genomic_offset/RDA/present_climatic_data_raster.Rdata")
#dataframe
write_xlsx(Present_climatic_data_T_adapcon_gentree_scaled ,path="C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/validation_GO/climatic_data/Present_climatic_data_T_adapcon_gentree_scaled.xlsx")
```

## Future data

### Data frame format

We finally can scale the future climatic data using the mean and sd of the past climatic data. 
```{r normalize and save the future data}
Future_climatic_data_T_adapcon_gentree_scaled <- data.frame(Future_climatic_data_T_adapcon_gentree[,c(1,2)],scale(Future_climatic_data_T_adapcon_gentree[,-c(1,2)], center = center_env_value_new_cli, scale = scale_env_value_new_cli))
```


```{r save df future data, include=FALSE}
#save df raw for GO using RDA because the genomic offset function is doing it using the center and sd values above
write_xlsx(Future_climatic_data_T_adapcon_gentree,"C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/Future_climatic_data_T_adapcon_gentree.xlsx")
save(Future_climatic_data_T_adapcon_gentree,file="C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/Future_climatic_data_T_adapcon_gentree.Rdata")

#save scale df
write_xlsx(Future_climatic_data_T_adapcon_gentree_scaled,"C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/Future_climatic_data_T_adapcon_gentree_scaled.xlsx")
save(Future_climatic_data_T_adapcon_gentree_scaled,file="C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/Future_climatic_data_T_adapcon_gentree_scaled.Rdata")

```

### Raster format

The raster needs to be the mean value of the 5 climatic models 
```{r create 1 list for each bio from each models, eval=F, echo=T}
library(raster)

# List of climatic variables
list_climatic_var <- c("bio1", "bio2","bio4", "bio9","bio12","bio15")  # Add more variables as needed

# List of climatic models
list_models <- c("GFDL_ESM4", "IPSL_CM6A_LR", "MPI_ESM1_2_HR", "MPI_ESM2_0", "UKESM1_0_LL")

# Create empty lists to store rasters for each variable
bio1_rasters <- list()
bio2_rasters <- list()
bio4_rasters <- list()
bio9_rasters <- list()
bio12_rasters <- list()
bio15_rasters <- list()

for (var_clim in list_climatic_var) {
  for (model in list_models) {
    # Create file path
    file_path <- paste0("C:/Users/tfrancisco/Documents/Thesis/Data/Climatic/Future_climate/2041_2070_", model, "_ssp370/", var_clim, ".tif")
    
    # Read raster
    raster_data <- raster(file_path)
    
    # Store raster in corresponding list based on variable
    if (var_clim == "bio1") {
      bio1_rasters[[model]] <- raster_data
    } else if (var_clim == "bio2") {
      bio2_rasters[[model]] <- raster_data
    } else if (var_clim == "bio4
               ") {
      bio4_rasters[[model]] <- raster_data
    } else if (var_clim == "bio9") {
      bio9_rasters[[model]] <- raster_data
    } else if(var_clim =="bio12"){
      bio12_rasters[[model]] <- raster_data
    } else if(var_clim =="bio15"){
      bio15_rasters[[model]] <- raster_data
    }
  }
}

```

```{r calculate mean raster, eval=F, echo=T}

calculate_mean_raster <- function(raster_list) {
  raster_stack <- stack(raster_list)
  mean_raster <- calc(raster_stack, mean)
  return(mean_raster)
}

list_raster_to_mean <- c("bio1_rasters","bio2_rasters","bio4_rasters","bio9_rasters","bio12_rasters","bio15_rasters")
name_file <- c("bio1","bio2","bio4","bio9","bio12","bio15")
for(x in 1:length(list_raster_to_mean)){
  
  var <- get(list_raster_to_mean[x])
  # Calculate mean raster for each climatic variable
mean_bio <- calculate_mean_raster(var)
name_file_use <- name_file[x]

assign(name_file_use,mean_bio)  
}
```

```{r save raster mean file, eval=F, echo=T}
for(x in 1:length(name_file)){
  var <- name_file[x]
  writeRaster(get(var), paste0("C:/Users/tfrancisco/Documents/Thesis/Data/Climatic/Mean_future_raster/",var), format="GTiff")
}
```

Now, we can load the folder with raster and create the raster stack with all the climatic variables

```{r create the file with six rasters and save file}
future_climatic_data_raster <- stack(list.files(paste0("C:/Users/tfrancisco/Documents/Thesis/Data/Climatic/Mean_future_raster/"), pattern = ".tif$", full.names = T)) 

past_climatic_data_raster <- data.frame(meta_data_pop$Population,raster::extract(future_climatic_data_raster, coords))#important to add raster:: because
#names of the raster (because they are not in the classic order Bio 1, 2 etc they are order by number so bio 1, 10, 11 etc)
names(future_climatic_data_raster) <- unlist(strsplit(unlist(lapply(strsplit(list.files(paste0("C:/Users/tfrancisco/Documents/Thesis/Data/Climatic/Mean_future_raster/"), pattern = ".tif$", full.names = T), split = paste0("C:/Users/tfrancisco/Documents/Thesis/Data/Climatic/Mean_future_raster/")), function(x) x[2])), split = ".tif"))

#save file
save(future_climatic_data_raster,file="C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Genomic_offset/RDA/future_climatic_data_raster.Rdata",F=T)
```


## LFMM analysis

For the LFMM analysis, we also need to create a data frame with climatic values at the individual level (where individuals of the same population have the same climatic values) because this analysis works at the individual level.

```{r new climatic matrix at individual-level}
#load meta_data at individual level

climatic_data_indivdual_level_new_selection <- merge(meta_data_vcf[,-2],past_climatic_data_order[,c(1:4,6,10,12,15)],"Population")

climatic_data_indivdual_level_scaled_new_selection <- climatic_data_indivdual_level_new_selection[,-c(1:3)] %>% 
  scale()

climatic_data_indivdual_level_scaled_new_selection_df <- data.frame(climatic_data_indivdual_level_new_selection[,c(1:3)],climatic_data_indivdual_level_scaled_new_selection)

#rename columns
colnames(climatic_data_indivdual_level_scaled_new_selection_df) <- c("Population","VCF_ID","Country","Annual_Tc","Diurnal_range_Tc","Tc_Seasonality","Tc_driest_quarter","Annual_P","P_Seasonality")
```

```{r save LFMM data, include=FALSE}
write_xlsx(climatic_data_indivdual_level_scaled_new_selection_df,"C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/climatic_data_indivdual_level_scaled_new_selection_df.xlsx")
save(climatic_data_indivdual_level_scaled_new_selection_df,file="C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/climatic_data_indivdual_level_scaled_new_selection_df.Rdata")

```





**What is below is a draft for future climatic data that might be usefull**

```{r data_to_be_used_raster future, eval=FALSE, include=FALSE}
Future_clim_data <- data.frame(merge(meta_data_pop[,c(2,4,5)],Future_climatic_data_T_adapcon_gentree[,c(1,3,7,8,4:6)],"Population"))
# we need to rename colnames for raster
Future_clim_data$Latitude <- as.numeric(Future_clim_data$Latitude);Future_clim_data$Longitude <- as.numeric(Future_clim_data$Longitude)
Future_clim_data_new <- data.frame(Future_clim_data[,c(3,2)],Future_clim_data[,c(4:9)])

colnames(Future_clim_data_new) <- c('x', 'y',"bio1", "bio12", "bio15","bio2","bio4","bio9")
```

```{r rasterization, eval=FALSE, include=FALSE}
# create a raster object with the extent of the data
r_obj <- raster(xmn=min(Future_clim_data_new$x)-0.1, xmx=max(Future_clim_data_new$x)+0.1, ymn=min(Future_clim_data_new$y)-0.1, ymx=max(Future_clim_data_new$y)+0.1, resolution=c(0.01,0.01)) #the resolution is smaller than what we have in meta_data_pop

# use rasterize to create desired raster with the pop and their values
future_climatic_data_raster <- rasterize(x=Future_clim_data_new[, 1:2], # lon-lat data
                    y=r_obj, # raster object
                    field=Future_clim_data_new[, 3:8], # clim var to fill raster with
                    fun=mean) # aggregate function
```

```{r save raster futut data, eval=FALSE, include=FALSE}
save(future_climatic_data_raster,file="C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Genomic_offset/RDA/future_climatic_data_raster.Rdata",F=T)
```


