---
title: "GF_genomic_offset"
author: "Thomas Francisco"
date: "2024-05-15"
output:
  html_document:
    number_sections: true #titles
    toc: true #table of content
    toc_float: true # enable the toc to be on the side of the text, always visible
    collapsed: True #control if the toc label will only display top level titles
    toc_depth: 3
---
```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache = FALSE
)
library(ggplot2)
library(gradientForest)
library(pdist)
library(corrplot)
library(dplyr)
library(rnaturalearth)
```

# Introduction 

This script aims to calculate the genomic offset using gradient forest. The genomic offset concept have already been explained in the script RDA_genomic_offset.
    
Here, the GEA relationship was estimated using the Gradient forest machine learning method (see explanations in the GF_candidate_selection script).

# Data

## Climatic data
      
```{r meta data}
meta_data_pop <- read.csv("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Populations/taxus_sample_29pop.csv",h=T,sep=";",dec=",")
#alphabetic order
meta_data_pop_order <- meta_data_pop[order(meta_data_pop$Population),]
```


Climatic data need to be in a dataframe:
```{r load Past climatic data}
#Past climatic data
past_climatic <- read.csv("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/Past_new_6_Climatic_data_scale_df.csv",sep=";",dec=",")
vars <- colnames(past_climatic[,-c(1:2)])
```

- Present climatic data: 
```{r load present climatic data}
#Present climatic data
present_climatic <- read.csv("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/validation_GO/climatic_data/Present_climatic_data_T_adapcon_gentree_scaled.csv",sep=";",dec=",")
```

- Future climatic data: 
```{r future clim data}
future_climatic <- read.csv("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/Future_climatic_data_T_adapcon_gentree_scaled.csv",sep=";",dec=",")
```
 
## Genomic data

```{r genomic data}
#genomic data
load("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/genetic_new/data_allelic_frequencies_29pop_adapcon_gentree_475_8616.Rdata")
genomic_matrix <- data_allelic_frequencies_29pop_adapcon_gentree_475_8616
```

## Set of SNPs

We can also load the sets of retained SNPs.   
We have several sets:  
- One with random SNPs  
- One with random SNPs with the same AF as the set of outliers (LC)  
- One with less conservative thresholds  
- One with more conservative thresholds  
- One with the LC SNPs in common with the genomic data of the clonal bank populations   
```{r load set of markers}
list_set <- c("random_neutral_set_SNPs_T_adapcon_gentree_bis","random_set_taxus","outliers_set_final_overlapping_no_LD_LC_new_var","outliers_set_final_overlapping_no_LD_new_var","CG_common_set")

for(i in 1:length(list_set)){
  
  set <- list_set[i]
  
  load(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/GEA_new_var/outliers/",set,".Rdata"))
}
list_random_SNPs <- colnames(random_neutral_set_SNPs_T_adapcon_gentree_bis)
list_random_SNPs_same_AF <- random_set_taxus$name_snps
```
      
# GEA relationship

The next step is to calculate the relationship between the outliers and climatic variables using the nonlinear model of Gradient Forest (GF). This follows the same principle as candidate selection with GF; however, here we will summarize the information in the turnover function for each SNP using all predictors to create a single turnover function for all response variables.

- Model with all the SNPs:
```{r GEA using gradient forest all snps}
#Run_GEA_GF_all <- gradientForest(data.frame(past_climatic[,vars],genomic_matrix), 
#                             predictor.vars=vars, 
#                             response.vars=colnames(genomic_matrix),
#                            corr.threshold=0.5, ntree=500, trace=T)

#length(Run_GEA_GF_all$result)
```

```{r save GEA GF random, include=FALSE}
#save(Run_GEA_GF_all,file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/GF/RUNs/Run_GEA_GF_all.Rdata")
```

- Other models:
```{r GEA using gradient forest}
#list_set <- c("list_random_SNPs","list_random_SNPs_same_AF","outliers_set_final_overlapping_no_LD_LC_new_var","outliers_set_final_overlapping_no_LD_new_var","CG_common_set")
#list_name <- c("random","random_same_AF","LC","MC","CG")

#for(i in 1:lenght(list_set)){
  
#  set <- get(list_set[i])
#  name <- list_name[i]
  
  #Run_GEA_GF <- gradientForest(data.frame(past_climatic[,vars],genomic_matrix[set]), 
#                             predictor.vars=vars, 
#                             response.vars=colnames(genomic_matrix[set]),
#                            corr.threshold=0.5, ntree=500, trace=T)

#length(Run_GEA_GF$result)
  
  #assign(paste0("Run_GEA_GF_",name),Run_GEA_GF)
#}
```


```{r save GEA GF}
#list_model <- c("all","random","random_same_AF","LC","MC","CG")

#for(i in 1:length(list_model)){
  
#  model <- list_model[i]
  
#  save(list = paste0("Run_GEA_GF_", model), file = paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/GF/RUNs/Run_GEA_GF_",model,".Rdata"))
#}
```

```{r load GEA GF}

list_model <- c("all","random","random_same_AF","LC","MC","CG")

for(i in 1:length(list_model)){
  
  model <- list_model[i]
  
load(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/GF/RUNs/Run_GEA_GF_",model,".Rdata"))
}
```

Example LC set:  We can see that 83 out of the 98 outliers are associated with some of the predictors. These 83 SNPs will be used by Gradient Forest (GF) in the next step to estimate the genomic offset.  

# Interpolate/extrapolate the GEA relationship and calculate genomic offset
      
Now that we have our GEA models, we can interpolate/extrapolate the relationship across space and time to estimate the past and future genomic composition for calculating the genomic offset metric. To do this, we modified the function from Matthew C. Fitzpatrick (github: https://github.com/fitzLab-AL/geneticOffsetR/blob/main/poplarGBS.gf.supportFunctions.R). 
      
## Function

```{r genomic offset function}
########## calculate adaptive offset for populations in space or time
genomic_offset_function <- function(gfMod, vars, env2, combined=F,
                       pops = envPop$pop_code, weighted=FALSE){
  
  #gfMod = gf model for prediction
  #vars = names of env variables
  #env2 = new environment (new place / time)
    transEnv2 <- predict(gfMod, env2[,vars]) #new env
    transEnv1 <- predict(gfMod) #current env
  
  #calculate Euclidean distance in transformed env space
  num <- nrow(transEnv1)
  dOut <- lapply(1:num, function(x, tEnv1, tEnv2){
    as.numeric(pdist(tEnv1[x,],  tEnv2[x,])@dist)}, tEnv2=transEnv2, tEnv1=transEnv1)
  return(dOut)
}
```

## Models

We can apply this function to our dataset: 
```{r Run genomic offset function}
list_set <- c("Run_GEA_GF_all","Run_GEA_GF_random","Run_GEA_GF_random_same_AF","Run_GEA_GF_LC","Run_GEA_GF_MC","Run_GEA_GF_CG")
list_period <- c("present", "future")

#vars = past climatic data used in GEA

for(i in 1: length(list_period)){
  
  period <- list_period[i]


for(x in 1:length(list_set)){
  GF_run <- get(list_set[x])
  name <- c("all","random","random_same_AF","LC","MC","CG")
  
  
  climate_to_calculate_GO <- get(paste0(period,"_climatic"))
  
Genomic_offset <- genomic_offset_function(gfMod=GF_run, vars=vars, env2=climate_to_calculate_GO[,vars], combined=F,
                       pops = row.names(genomic_matrix), weighted=FALSE)

#extraction GO values
Genomic_offset$values <- unlist(Genomic_offset)

genomic_offset_GF <- data.frame(Population=row.names(genomic_matrix),GO=Genomic_offset$values)
names(genomic_offset_GF)[2] <- paste0("genomic_offset_GF_",name[x])

assign(paste0("genomic_offset_GF_",name[x],"_",period),genomic_offset_GF)
  }
}
```

```{r save genomic offset values, include=FALSE}
list_period <- c("present","future")
list_set <- c("all","random","random_same_AF","LC","MC","CG")

for(i in 1:length(list_set)){
  set <- list_set[i]
  for(x in 1:length(list_period)){
    period <- list_period[x]
    name_file <- paste0(set,"_",period)
    
  file_save<- get(paste0("genomic_offset_GF_",set,"_",period))
  name <- paste0("GO_T_Adapcon_Gentree_GF_",name_file)
  assign(name,file_save)
  save(list = name,file=paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/GF/data/",name,".Rdata"))
  }
}
```

## Graphical representations

Based on these genomic offset values, we can create similar plots to the ones for RDA GO, displaying the values for each population.
```{r plot GO values,echo=T,eval=F}
list_period <- c("present","future")
list_set <- c("all","random","random_same_AF","LC","MC","CG")

for(i in 1:length(list_period)){
  
  period <- list_period[i]
  
  for(x in 1: length(list_set)){
    
    set <- list_set[x]
GO_df <- get(paste0("GO_T_Adapcon_Gentree_GF_",set,"_",period))
    
     #first, we need to add the coordinates
Genomic_offset_coord <- merge(GO_df,meta_data_pop_order[,c(2,4,5)],"Population")

#transform longitude and latitude to numeric variables
Genomic_offset_coord <- Genomic_offset_coord %>% mutate(Longitude=as.numeric(Longitude),Latitude=as.numeric(Latitude))


colors <- c( "darkgreen", "#FDF7B3","#FC4E2A","#BD0026","darkorchid4")
#background map
admin <- ne_countries(scale = "medium", returnclass = "sf")
      
plot <- ggplot(data = Genomic_offset_coord) + 
  geom_sf(data = admin, fill = gray(0.92), size = 0) +
  geom_point(aes(x = Longitude, y = Latitude, fill = cut_number(Genomic_offset_coord[,2], n = 5)), shape = 21,size=3, color = "black") +
  scale_fill_manual(
    values = colors,
    labels = c("low values","","","","high values"),
    drop = FALSE,na.translate = FALSE)+  # Ensure all levels are shown in the legend
  geom_sf(data = admin, fill = NA, size = 0.1) +
  coord_sf(xlim = c(-10, 30), ylim = c(35, 62), expand = FALSE) +
  xlab("Longitude") + ylab("Latitude") +
  guides(fill = guide_legend(title = "Genomic offset")) +
  ggtitle(paste0("Genomic offset across populations ",set," ",period)) +
  theme_bw(base_size = 11) +
  theme(legend.position="right", panel.grid = element_blank(), strip.text = element_text(size=11),plot.title = element_text(hjust = 0.5,color = "Black",face="italic"))
      
print(plot) 

#save
     pdf(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/GF/figures/GO/Genomic_offset_across_populations_",set,"_",period,".pdf"));print(plot);dev.off()
  }
}
```

```{r graph rmarkdown 1, echo=FALSE}
list_period_rmarkdown <- c("present","future")
list_set_rmarkdown <- c("LC")

for(i in 1:length(list_period_rmarkdown)){
  period <- list_period_rmarkdown[i]
  for(x in 1: length(list_set_rmarkdown)){
    set <- list_set_rmarkdown[x]
GO_df <- get(paste0("GO_T_Adapcon_Gentree_GF_",set,"_",period))
     #first, we need to add the coordinates
Genomic_offset_coord <- merge(GO_df,meta_data_pop_order[,c(2,4,5)],"Population")
#transform longitude and latitude to numeric variables
Genomic_offset_coord <- Genomic_offset_coord %>% mutate(Longitude=as.numeric(Longitude),Latitude=as.numeric(Latitude))
colors <- c( "darkgreen", "#FDF7B3","#FC4E2A","#BD0026","darkorchid4")
#background map
admin <- ne_countries(scale = "medium", returnclass = "sf")
plot <- ggplot(data = Genomic_offset_coord) + 
  geom_sf(data = admin, fill = gray(0.92), size = 0) +
  geom_point(aes(x = Longitude, y = Latitude, fill = cut_number(Genomic_offset_coord[,2], n = 5)), shape = 21,size=3, color = "black") +
  scale_fill_manual(
    values = colors,
    labels = c("low values","","","","high values"),
    drop = FALSE,na.translate = FALSE)+  # Ensure all levels are shown in the legend
  geom_sf(data = admin, fill = NA, size = 0.1) +
  coord_sf(xlim = c(-10, 30), ylim = c(35, 62), expand = FALSE) +
  xlab("Longitude") + ylab("Latitude") +
  guides(fill = guide_legend(title = "Genomic offset")) +
  ggtitle(paste0("Genomic offset across populations ",set," ",period)) +
  theme_bw(base_size = 11) +
  theme(legend.position="right", panel.grid = element_blank(), strip.text = element_text(size=11),plot.title = element_text(hjust = 0.5,color = "Black",face="italic"))
print(plot) 
  }
}
```

## Correlation across genomic offset prediction  

We observed that there is no clear pattern in the genomic offset; it does not correlate with continentality, coastal proximity, or other discernible patterns. We hypothesized that elevation might influence genomic offset, with populations at higher altitudes potentially exhibiting higher genomic offsets.

```{r correlation GO elevation}
#load elevation

elevation_data <- read.csv("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/ClimateDT_extraction/extraction_climateDT_29pop_T.csv",h=T,sep=";",dec=",")

#merge elevation data with GO
data_GO_altitude <- data.frame(merge(genomic_offset_GF_LC_present,elevation_data,"Population"))
data_GO_altitude_df <- data_GO_altitude %>% 
  mutate(Longitude=as.numeric(Longitude),Latitude=as.numeric(Latitude),elevation=as.numeric(Elevation.DEM_90m.))

#df
df_tot_GO_predictions_present <- data.frame(Population=genomic_offset_GF_LC_future[,1],GO_all_pres=genomic_offset_GF_all_present[,2],GO_rand_pres=genomic_offset_GF_random_present[,2],GO_rand_same_AF_pres=genomic_offset_GF_random_same_AF_present[,2],GO_LC_pres=genomic_offset_GF_LC_present[,2],GO_MC_pres=genomic_offset_GF_MC_present[,2],GO_CG_pres=genomic_offset_GF_CG_present[,2],elevation=data_GO_altitude_df$elevation)

correlation_present <- cor(df_tot_GO_predictions_present[,-1])
corrplot(correlation_present, method = "number", addrect = 2, col = c("darkorange","darkred"), type = "lower", tl.col = "black", tl.cex = 0.6, number.cex = 0.6)

df_tot_GO_predictions_future <- data.frame(Population=genomic_offset_GF_LC_future[,1],GO_all_Fut=genomic_offset_GF_all_future[,2],GO_rand_Fut=genomic_offset_GF_random_future[,2],GO_rand_same_AF_Fut=genomic_offset_GF_random_same_AF_future[,2],GO_LC_Fut=genomic_offset_GF_LC_future[,2],GO_MC_Fut=genomic_offset_GF_MC_future[,2],GO_CG_Fut=genomic_offset_GF_CG_future[,2],elevation=data_GO_altitude_df$elevation)

correlation_future <- cor(df_tot_GO_predictions_future[,-1])
corrplot(correlation_future, method = "number", addrect = 2, col = c("darkorange","darkred"), type = "lower", tl.col = "black", tl.cex = 0.6, number.cex = 0.6)
```

```{r save corrplot}
pdf("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/GF/figures/comparison/correlation_GO_present_values_GF_T_set_period.pdf");corrplot(correlation_present, method = "number", addrect = 2, col = c("darkorange","darkred"), type = "lower", tl.col = "black", tl.cex = 0.6, number.cex = 0.6);dev.off()

pdf("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/GF/figures/comparison/correlation_GO_future_values_GF_T_set_period.pdf");corrplot(correlation_future, method = "number", addrect = 2, col = c("darkorange","darkred"), type = "lower", tl.col = "black", tl.cex = 0.6, number.cex = 0.6);dev.off()
```
We observed a positive correlation between genomic offset and elevation, which vary depending on the set of SNPs used.  

More generally, we can see the same results as for RDA for the correlation between models and periods. 
For the MC and LC sets, we have almost the same GO values regardless of the period. 
For the comparison neutral vs outliers set, we have also similar results with an overall correlation of 0.77 and 0.76.  
We can investigate the rank of populations according to their genomic offset values to further investigate these findings.

## Rank 

```{r Rank of outliers populations,echo=T,eval=F}
list_period <- c("present","future")
list_set <- c("all","random","random_same_AF","LC","MC","CG")

for(i in 1:length(list_period)){
  
  period <- list_period[i]
  
  for(x in 1:(length(list_set)-1)){
    
    set <- list_set[x]
    set2 <- list_set[x+1]
    
    #merge two dataset
GO_RDA_merge <- merge(GO_df <- get(paste0("genomic_offset_GF_",set,"_",period)),GO_df <- get(paste0("genomic_offset_GF_",set2,"_",period)),"Population")

#df
GO_RDA_set_df<- data.frame(Population= GO_RDA_merge$Population,GO_RDA_1=GO_RDA_merge[,2],GO_RDA_2=GO_RDA_merge[,3])

#RDA_GO rank
GO_RDA_set_df$rank_1 <- rank(GO_RDA_set_df$GO_RDA_1)

#GF_GO rank
GO_RDA_set_df$rank_2 <- rank(GO_RDA_set_df$GO_RDA_2)


#plot
GO_RDA_set_df_meta <- merge(GO_RDA_set_df,meta_data_pop_order[,c(1,2)],"Population")

GO_RDA_set_df_meta$Country <- as.factor(GO_RDA_set_df_meta$Country)

Scatterplot <- ggplot(GO_RDA_set_df_meta,aes(x=rank_1,y=rank_2)) +
  geom_point(aes(color=Country),size=3)+
  scale_colour_manual(name="Countries",
                      values = c("orangered3","gold2","darkorchid3","navyblue","turquoise2","green3","blue","red","black","gray","orange","darkgreen")) +
  geom_abline(intercept = 0, slope = 1, color="gray60")+
    ggtitle(paste0("Comparison GO rank of populations RDA ",set,"/",set2," ",period))+ 
    theme_set(theme_bw())

plot(Scatterplot)


pdf(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/GF/figures/comparison/Scatterplot_comparison_rank_pop_",set,"_",set2,"_",period,"_Taxus.pdf"));print(Scatterplot);dev.off()
  }
}
```

```{r graph rmarkdown 2, echo=FALSE}
list_period_rmarkdown <- c("present","future")
list_set_rmarkdown <- c("LC","MC")

for(i in 1:length(list_period_rmarkdown)){
  period <- list_period_rmarkdown[i]
  for(x in 1:(length(list_set_rmarkdown)-1)){
    set <- list_set_rmarkdown[x]
    set2 <- list_set_rmarkdown[x+1]
    #merge two dataset
GO_RDA_merge <- merge(GO_df <- get(paste0("genomic_offset_GF_",set,"_",period)),GO_df <- get(paste0("genomic_offset_GF_",set2,"_",period)),"Population")
#df
GO_RDA_set_df<- data.frame(Population= GO_RDA_merge$Population,GO_RDA_1=GO_RDA_merge[,2],GO_RDA_2=GO_RDA_merge[,3])
#RDA_GO rank
GO_RDA_set_df$rank_1 <- rank(GO_RDA_set_df$GO_RDA_1)
#GF_GO rank
GO_RDA_set_df$rank_2 <- rank(GO_RDA_set_df$GO_RDA_2)
#plot
GO_RDA_set_df_meta <- merge(GO_RDA_set_df,meta_data_pop_order[,c(1,2)],"Population")
GO_RDA_set_df_meta$Country <- as.factor(GO_RDA_set_df_meta$Country)
Scatterplot <- ggplot(GO_RDA_set_df_meta,aes(x=rank_1,y=rank_2)) +
  geom_point(aes(color=Country),size=3)+
  scale_colour_manual(name="Countries",
                      values = c("orangered3","gold2","darkorchid3","navyblue","turquoise2","green3","blue","red","black","gray","orange","darkgreen")) +
  geom_abline(intercept = 0, slope = 1, color="gray60")+
    ggtitle(paste0("Comparison GO rank of populations RDA ",set,"/",set2," ",period))+ 
    theme_set(theme_bw())
plot(Scatterplot)
  }
}
```


## Temporal trend

We can also examine the trend of maladaptation to determine whether populations currently exhibiting higher predicted maladaptation will continue to show the same pattern in the future.

```{r trend GO, echo=T,eval=F}
list_set <- c("all","random","random_same_AF","LC","MC","CG")

for(i in 1:length(list_set)){
  
  set <- list_set[i]
  
  period_1 <- c("present")
  period_2 <- c("future")
  
  df_pres <- get(paste0("genomic_offset_GF_",set,"_",period_1))
  df_fut <- get(paste0("genomic_offset_GF_",set,"_",period_2))
  
  #df
df_f_pres <- data.frame(Population = df_pres[,1], Values = df_pres[,2])
df_f_fut <- data.frame(Population = df_fut[,1], Values = df_fut[,2])

#add period
df_f_pres$Period <- "Present"
df_f_fut$Period <- "Future"


#combine them
df_GO <- rbind(df_f_fut,df_f_pres)
df_GO$Period <- factor(df_GO$Period, levels = c("Present", "Future"))

#plot

trend_plot <- ggplot(df_GO, aes(x = Period, y = Values, group = Population)) +
  geom_line(aes(color = Population)) +   # Line plot, connecting the points for each population
  geom_point(aes(color = Population)) +  # Adding points for each value
  theme_minimal() +
  labs(title = paste0("Genomic offset GF trend across present and future ",set),
       x = "Period",
       y = "Genomic offset prediction") +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 11),
    plot.title = element_text(hjust = 0.5, color = "Black", face = "italic")
    )
    plot(trend_plot)
#save
    pdf(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/RDA/figures/comparison/GO_pres_future/Scatterplot_comparison_trend_GO_present_future_",set,"_Taxus.pdf"));print(trend_plot);dev.off()
}
```

```{r graph rmarkdown 3, echo=FALSE}
list_set_rmarkdown <- c("LC")

for(i in 1:length(list_set_rmarkdown)){
  set <- list_set_rmarkdown[i]
  period_1 <- c("present")
  period_2 <- c("future")
  df_pres <- get(paste0("genomic_offset_GF_",set,"_",period_1))
  df_fut <- get(paste0("genomic_offset_GF_",set,"_",period_2))
  #df
df_f_pres <- data.frame(Population = df_pres[,1], Values = df_pres[,2])
df_f_fut <- data.frame(Population = df_fut[,1], Values = df_fut[,2])
#add period
df_f_pres$Period <- "Present"
df_f_fut$Period <- "Future"
#combine them
df_GO <- rbind(df_f_fut,df_f_pres)
df_GO$Period <- factor(df_GO$Period, levels = c("Present", "Future"))
#plot
trend_plot <- ggplot(df_GO, aes(x = Period, y = Values, group = Population)) +
  geom_line(aes(color = Population)) +   # Line plot, connecting the points for each population
  geom_point(aes(color = Population)) +  # Adding points for each value
  theme_minimal() +
  labs(title = paste0("Genomic offset GF trend across present and future ",set),
       x = "Period",
       y = "Genomic offset prediction") +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 11),
    plot.title = element_text(hjust = 0.5, color = "Black", face = "italic")
    )
    plot(trend_plot)
}
```

# Global conclusion

We observed that the LC and MC sets are nearly identical; therefore, we will only perform further analyses the analyses on the LC set and not on the MC set.


# Draft

Example of GEA method 1 model: 
```{r GEA using gradient forest MC, eval=FALSE, include=FALSE}

#Run_GEA_GF_MC <- gradientForest(data.frame(past_climatic[,vars],genomic_matrix[outliers_set_final_overlapping_no_LD_new_var]), 
#                             predictor.vars=vars, response.vars=colnames(genomic_matrix[outliers_set_final_overlapping_no_LD_new_var]),
#                             corr.threshold=0.5, ntree=500, trace=T)

#length(Run_GEA_GF_MC$result)

#save(Run_GEA_GF_MC,file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/GF/RUNs/Run_GEA_GF_MC.Rdata")
#load("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/GF/RUNs/Run_GEA_GF_MC.Rdata")
```