---
title: "Phenotypic_data_clonal_bank"
author: "Thomas Francisco"
date: "2024-09-16"
output:
  html_document:
    number_sections: true #titles
    toc: true #table of content
    toc_float: true # enable the toc to be on the side of the text, always visible
    collapsed: True #control if the toc label will only display top level titles
    toc_depth: 3
---

```{r library, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
rm(list = ls()) 
library(MCMCglmm)
library(lme4)
library(coda)
library(parallel)
library(dplyr)
library(stargazer)
library(ggplot2)
library(pastecs)
library(readxl)
library(purrr)
library(QGglmm)
library(purrr)
library(coda)
library(writexl)
```

# Introduction

The goal of this script is to estimate the BLUP values for the phenotypic data that will be used to evaluate the genomic offset predictions. BLUP estimates allow us to generate phenotypic population values for each trait while accounting for covariates, such as genotype for clones, tree ID when we have multiple measurements per tree, and tree size to account for age in growth-related traits.
Depending on the trait, we will select a specific period for climate data extraction. For each trait, the initial year is 1992 (the year of establishment of the clonal bank), with the following end years:   
- 2021 for leaf thickness  
- 2012 for shoot volume, stem length, spring elongation, and total open male strobili  

# Phenotypic data

We loaded the phenotypic traits associated with: 

Growth -> Shoot volume
Growth_pheno -> spring elongation
Repro_pheno -> total open strobili
leaf_traits -> leaf thickness

The first step is to load the functions that was used in the analyses. 
```{r function}
clean.MCMC <- function(x) {
  sols <- summary(x)$solutions  ## pull out relevant info from model summary
  Gcovs <- summary(x)$Gcovariances
  Rcovs <- summary(x)$Rcovariances
  fixed <- data.frame(row.names(sols), sols, row.names = NULL)  ## convert to dataframes with the row.names as the first col
  random <- data.frame(row.names(Gcovs), Gcovs, row.names = NULL)
  residual <- data.frame(row.names(Rcovs), Rcovs, row.names = NULL)
  names(fixed)[names(fixed) == "row.names.sols."] <- "variable"  ## change the columns names to variable, so they all match
  names(random)[names(random) == "row.names.Gcovs."] <- "variable"
  names(residual)[names(residual) == "row.names.Rcovs."] <- "variable"
  fixed$effect <- "fixed"  ## add ID column for type of effect (fixed, random, residual)
  random$effect <- "random"
  residual$effect <- "residual"
  modelTerms <- as.data.frame(bind_rows(fixed, random, residual))  # merge it all together
}


plot.estimates <- function(x) {
  if (class(x) != "summary.mcmc")
    x <- summary(x)
  n <- dim(x$statistics)[1]
  par(mar=c(2, 9, 4, 1))
  plot(x$statistics[,1], n:1,
       yaxt="n", ylab="",
       xlim=range(x$quantiles)*1.2,
       pch=19,
       main="Posterior means and 95% credible intervals")
  grid()
  axis(2, at=n:1, rownames(x$statistics), las=2)
  arrows(x$quantiles[,1], n:1, x$quantiles[,5], n:1, code=0)
  abline(v=0, lty=2)
}


myFun_h2Qst1 <- function(mcmc){  
  out <- c(
    posterior.mode((mcmc$VCV[,2])/(rowSums(mcmc$VCV))),
    HPDinterval((mcmc$VCV[,2])/(rowSums(mcmc$VCV)))[1],
    HPDinterval((mcmc$VCV[,2])/(rowSums(mcmc$VCV)))[2],
    posterior.mode((mcmc$VCV[,2])/((mcmc$VCV[,2])+(mcmc$VCV[,3]))),
    HPDinterval((mcmc$VCV[,2])/((mcmc$VCV[,2])+(mcmc$VCV[,3])))[1],
    HPDinterval((mcmc$VCV[,2])/((mcmc$VCV[,2])+(mcmc$VCV[,3])))[2],
    posterior.mode((mcmc$VCV[,1])/(rowSums(mcmc$VCV))),
    HPDinterval((mcmc$VCV[,1])/(rowSums(mcmc$VCV)))[1],
    HPDinterval((mcmc$VCV[,1])/(rowSums(mcmc$VCV)))[2],
    posterior.mode((mcmc$VCV[,1])/(2*(mcmc$VCV[,2])+(mcmc$VCV[,1]))),
    HPDinterval((mcmc$VCV[,1])/(2*(mcmc$VCV[,2])+(mcmc$VCV[,1])))[1],
    HPDinterval((mcmc$VCV[,1])/(2*(mcmc$VCV[,2])+(mcmc$VCV[,1])))[2]
  )
  names(out) <- c("H2c_all","lo.ci.H2c_all","up.ci.H2c_all" ,"H2c","lo.ci.H2c","up.ci.H2c" ,"H2p","lo.ci.H2p","up.ci.H2p" ,"Qst","lo.ci.Qst","up.ci.Qst")
  return(out)}

# for the logit link, which is used when family=categorical
myFun_h2Qst2 <- function(mcmc){  
  out <- c(
    posterior.mode((mcmc$VCV[,2])/(rowSums(mcmc$VCV)+(pi^2)/3)),
    HPDinterval((mcmc$VCV[,2])/(rowSums(mcmc$VCV)+(pi^2)/3))[1],
    HPDinterval((mcmc$VCV[,2])/(rowSums(mcmc$VCV)+(pi^2)/3))[2],
    posterior.mode((mcmc$VCV[,2])/((mcmc$VCV[,2])+(mcmc$VCV[,3])+(pi^2)/3)),
    HPDinterval((mcmc$VCV[,2])/((mcmc$VCV[,2])+(mcmc$VCV[,3])+(pi^2)/3))[1],
    HPDinterval((mcmc$VCV[,2])/((mcmc$VCV[,2])+(mcmc$VCV[,3])+(pi^2)/3))[2],
    posterior.mode(((mcmc$VCV[,1]))/(rowSums(mcmc$VCV)+(pi^2)/3)),
    HPDinterval(((mcmc$VCV[,1]))/(rowSums(mcmc$VCV)+(pi^2)/3))[1],
    HPDinterval(((mcmc$VCV[,1]))/(rowSums(mcmc$VCV)+(pi^2)/3))[2],
    posterior.mode((mcmc$VCV[,1])/(2*(mcmc$VCV[,2])+(mcmc$VCV[,1]))),
    HPDinterval((mcmc$VCV[,1])/(2*(mcmc$VCV[,2])+(mcmc$VCV[,1])))[1],
    HPDinterval((mcmc$VCV[,1])/(2*(mcmc$VCV[,2])+(mcmc$VCV[,1])))[2]
  )
  names(out) <- c("H2c_all","lo.ci.H2c_all","up.ci.H2c_all" ,"H2c","lo.ci.H2c","up.ci.H2c" ,"H2pc","lo.ci.H2pc","up.ci.H2pc" ,"Qst","lo.ci.Qst","up.ci.Qst")
  return(out)}
```

## data
```{r phenotypic data}
list_traits <- c("Shoot_Growth","Basal_area","Shoot_Elongation","Male_Strobili","leaf_traits")
list_names <- c("growth","growth_bis","growth_pheno","repro_pheno","leaf_traits")

for(x in 1: length(list_traits)){
  
  trait <- list_traits[x]
  name <- list_names[x]
  
  path <- paste0("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/phenotypic_data/common_garden/Individual_level_data/Taxus_",trait,".xlsx")
phenotypic_data <- read_excel(path, na="NA")

assign(paste0("phenotypic_data_",name),phenotypic_data)
}
```

## Selection of the period

For each traits, we calculated average values across years as traits values in a given year are influenced by conditions and events from previous years. We retained the maximum number of sampling years while also maximizing the number of trees. 
```{r Growth, message=FALSE, warning=FALSE}
#Growth

##Shootvol
#identify the number of measurement per plant
phenotypic_data_growth_table <- data.frame(table(phenotypic_data_growth$Plant))

#filter the number of measurement to keep the maximum number of individuals
phenotypic_data_growth_filtered <- phenotypic_data_growth %>%
  filter(Plant %in% phenotypic_data_growth_table$Var1[phenotypic_data_growth_table$Freq == 4])

#mean values of the traits per plant
mean_traits_shootvol <- phenotypic_data_growth_filtered %>%
  group_by(Plant) %>%
  summarize(mean_AverageShootVolume = mean(AverageShootVolume.mm3, na.rm = TRUE),
            mean_StemLength = mean(StemLength.cm, na.rm = TRUE))

mean_traits_shootvol_merge <- merge(mean_traits_shootvol,phenotypic_data_growth,"Plant")

#keep only one value per plant (ex: 2012, it's not only the values of 2012)
mean_traits_growth_f <- mean_traits_shootvol_merge %>%
  filter(Year == 2012)

##Basal area
mean_traits_growth_bis_f <- data.frame(phenotypic_data_growth_bis) %>%
  filter(across(c(SumAB.2021.cm2, StemLength.cm), ~ !is.na(.)))
```

```{r Growth phenology}
# Growth phenology

##Spring elong
#identify the number of measurement per plant
phenotypic_data_phenology_table <- data.frame(table(phenotypic_data_growth_pheno$Plant))

#filter the number of measurement to keep the maximum number of individuals
phenotypic_data_pheno_filtered <- phenotypic_data_growth_pheno %>%
  filter(Plant %in% phenotypic_data_phenology_table$Var1[phenotypic_data_phenology_table$Freq == 3])

#mean values of the traits per plant
mean_traits_pheno <- phenotypic_data_pheno_filtered %>%
  group_by(Plant) %>%
  summarize(mean_springelong = mean(ProportionSpringElongation, na.rm = TRUE))

mean_traits_pheno_merge <- merge(mean_traits_pheno,phenotypic_data_growth_pheno,"Plant")

#keep only one value per plant (ex: 2012, it's not only the values of 2012)
mean_traits_growth_pheno_f <- mean_traits_pheno_merge %>%
  filter(Year == 2012)
```

```{r Reproductive phenology}
# Reproductive phenology

##Reproduction
#identify the number of measurement per plant
phenotypic_data_repro_table <- data.frame(table(phenotypic_data_repro_pheno$Plant))

#filter the number of measurement to keep the maximum number of individuals
phenotypic_data_repro_filtered <- phenotypic_data_repro_pheno %>%
  filter(Plant %in% phenotypic_data_repro_table$Var1[phenotypic_data_repro_table$Freq == 3])

#mean values of the traits per plant
mean_traits_repro <- phenotypic_data_repro_filtered %>%
  group_by(Plant) %>%
  summarize(mean_proportion_open = mean(Proportion_total_open, na.rm = TRUE))

mean_traits_repro_merge <- merge(mean_traits_repro,phenotypic_data_repro_pheno,"Plant")

#keep only one value per plant (ex: 2012, it's not only the values of 2012)
mean_traits_repro_pheno_f <- mean_traits_repro_merge %>%
  filter(Year == 2012)
```

```{r Leaf_traits}
#Leaf_traits
phenotypic_data_leaftrait_table <- data.frame(table(phenotypic_data_leaf_traits$Plant))
#table(phenotypic_data_leaftrait_table$Freq)#we retained 4

#filter the number of measurement to keep the maximum number of individuals
phenotypic_data_leaftrait_filtered <- phenotypic_data_leaf_traits %>%
  filter(Plant %in% phenotypic_data_leaftrait_table$Var1[phenotypic_data_leaftrait_table$Freq == 4])

#mean values of the traits per plant
mean_traits_leaftraits <- phenotypic_data_leaftrait_filtered %>%
  group_by(Plant) %>%
  summarize(mean_StomDens = mean(StomDens.mm2, na.rm = TRUE),
            mean_leafthickness = mean(LeafThickness.mm, na.rm = TRUE),
            mean_LMA = mean(LMA.gm2,na.rm=T),
            mean_D13C = mean(D13C,na.rm=T))

mean_traits_leaftraits_merge <- merge(mean_traits_leaftraits,phenotypic_data_leaf_traits[,c(1:5)],"Plant")

#keep only one value per plant (ex: 2012, it's not only the values of 2012)
mean_traits_leaftraits_f <- mean_traits_leaftraits_merge %>%
  filter(Year == "Y2021")
```

## Selection of the populations with enought trees

We also wanted to check the number of measured individuals per population, retaining only populations with more than two individuals
```{r number of measured indiv per pop, eval=F,echo=T}
table(mean_traits_growth_f$Population)
#We remove the ACEBDO, LASENIA

table(mean_traits_growth_pheno_f$Population)
#We remove the ACEBDO, ALCOY, LASENIA Quesada pop

table(mean_traits_repro_pheno_f$Population)
#We remove the BOCAHUERGANO, JERTE, LASENIA,PUEBLALILLO and RASQUERA 

table(mean_traits_leaftraits_f$Population)
#we can remove: ACEBEDO, ALCOY, LASENIA, and QUESADA
```
## Graphical visualization

To select the type of distribution we used in the GLMM models, we investigated the shape of the data distribution for each traits
```{r phenotypic_df_config}

list_traits <- c("growth","growth_bis","growth_pheno","repro_pheno","leaftraits")
for(x in 1: length(list_traits)){
  
  name <- list_traits[x]
  df <- get(paste0("mean_traits_",name,"_f"))
  
df$Population<- as.factor(df$Population)
df$Plant <- as.factor(df$Plant)
df$Genet <- as.factor(df$Genet)

assign(paste0("phenotypic_data_",name),df)
}
```

```{r hist}
hist(phenotypic_data_growth$mean_AverageShootVolume)
#hist(phenotypic_data_growth_bis$SumAB.2021.cm2)
hist(phenotypic_data_growth_pheno$mean_springelong)
hist(phenotypic_data_repro_pheno$mean_proportion_open)
#hist(phenotypic_data_leaftraits$mean_StomDens)
hist(phenotypic_data_leaftraits$mean_leafthickness)
#hist(phenotypic_data_leaftraits$mean_LMA)
#hist(phenotypic_data_leaftraits$mean_D13C)
```

# MCMC GLMM models
    
We can now run the MCMC GLMM models for each of the traits. We need to specify the priors as we are working with Bayesian models. We chose a non-informative prior due to the lack of prior information. We also set the number of iterations to 1,500,000, the burn-in period to 50,000, and the thinning interval to 500, sampling every 500 iterations. We ran each model four times per trait (4 MCMC chains) to ensure stability. All traits follow a Gaussian distribution with continuous data.

## Trait related to growth
    
### Shoot Volume

For growth traits, as trees in the clonal bank vary in age, we used stem length as a covariate to account for differences in shoot volume associated with older trees. 

```{r shootvolume model,echo=T,eval=F}
set.seed(3)

#phenotypic_data_growth$Sex <- as.factor(phenotypic_data_growth$Sex)
#phenotypic_data_growth_wo_Na <- phenotypic_data_growth %>%group_by(Sex) %>% drop_na()

#set the prior info for the fixed variables: year and the randoms ones: Population and Plant
prior_ShootVol=list(R=list(V=1, nu=0.002),G=list((G1=list(V=1, nu=0.002)),(G2=list(V=1, nu=0.002)),(G3=list(V=1, nu=0.002))))

#we can perform the model
ShootVol<- mclapply(1:4, function(i) {#we can perform 4 MCMC 
  ShootVol<-MCMCglmm(mean_AverageShootVolume~Sex,random=~Population + Genet +mean_StemLength,prior=prior_ShootVol, family="gaussian", nitt=1500000, burnin=50000, thin=500, data=phenotypic_data_growth_wo_Na, pr=TRUE, verbose=F) #the pheno is explain by the year as fixed because few number of parameters and by the Population and Plant factors, we are following a gaussian law due to the quantitative continuous type of the response variable (the pheno)
}, mc.cores=1)

#save the chains
growth_model_wo_plantID<- ShootVol
#save(ShootVol_model_wo_plantID,file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/ShootVol_model_wo_plantID.rda")
```

```{r load the MCMC growth }
load("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/ShootVol_model_wo_plantID.rda")
growth_model_wo_plantID <- ShootVol_model_wo_plantID
```

## Trait related to growth phenology

### Spring elongation
```{r springelong model,echo=T,eval=F}
set.seed(3)
#phenotypic_data_growth_pheno_wo_Na <- phenotypic_data_growth_pheno %>%group_by(Sex) %>% drop_na()
#phenotypic_data_growth_pheno_wo_Na$Sex <- as.factor(phenotypic_data_growth_pheno_wo_Na$Sex)

#set the prior info for the fixed variables: year and the randoms ones: Population and Plant
prior_springelong=list(R=list(V=1, nu=0.002),G=list((G1=list(V=1, nu=0.002)),(G2=list(V=1, nu=0.002))))

#we can perform the model

springelong<- mclapply(1:4, function(i) {#we can perform 4 MCMC 
  springelong<-MCMCglmm(mean_springelong~Sex,random=~Population + Genet,prior=prior_springelong, family="gaussian", nitt=1500000, burnin=50000, thin=500, data=phenotypic_data_growth_pheno_wo_Na, pr=TRUE, verbose=F) #the pheno is explain by the year as fixed because few number of parameters and by the Population and Plant factors, we are following a gaussian law due to the quantitative continuous type of the response variable (the pheno)
}, mc.cores=1)

#save the chains
growth_pheno_model_wo_plantID<- springelong
#save(springelong_model_wo_plantID,file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/springelong_model_wo_plantID.rda")
```

```{r load the MCMC growth pheno}
load("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/springelong_model_wo_plantID.rda")
growth_pheno_model_wo_plantID <- springelong_model_wo_plantID
```

## Trait related to reproductive phenology

### Proportion open strobili
```{r mature Strobili model,echo=T,eval=F}
set.seed(3)
#set the prior info for the fixed variables: year and the randoms ones: Population and Plant
prior_openstrob=list(R=list(V=1, nu=0.002),G=list((G1=list(V=1, nu=0.002)),(G2=list(V=1, nu=0.002))))

#we can perform the model

openstrob<- mclapply(1:4, function(i) {#we can perform 4 MCMC 
  openstrob<-MCMCglmm(mean_proportion_open~1,random=~Population + Genet,prior=prior_openstrob, family="gaussian", nitt=1500000, burnin=50000, thin=500, data=phenotypic_data_repro_pheno, pr=TRUE, verbose=F) #the pheno is explain by the year as fixed because few number of parameters and by the Population and Plant factors, we are following a poisson law due to the count type of the response variable (the pheno)
}, mc.cores=1)

#save the chains
openstrob_model_wo_plantID<- openstrob
#save(openstrob_model_wo_plantID,file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/openstrob_model_wo_plantID.rda")
```

```{r load the MCMC open strob}
load("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/openstrob_model_wo_plantID.rda")
repro_pheno_model_wo_plantID <- openstrob_model_wo_plantID
```

## Water statues 

### Leaf thickness
```{r Leaf thickness model,echo=T,eval=F}
set.seed(3)
#set the prior info for the fixed variables: year and the randoms ones: Population and Plant
prior_leaftraits=list(R=list(V=1, nu=0.002),G=list((G1=list(V=1, nu=0.002)),(G2=list(V=1, nu=0.002))))

#phenotypic_data_leaftraits$Sex<-as.factor(phenotypic_data_leaftraits$Sex)
#we can perform the model
leafthickness<- mclapply(1:4, function(i) {#we can perform 4 MCMC 
  leafthickness<-MCMCglmm(mean_leafthickness~1,random=~Population + Genet,prior=prior_leaftraits, family="gaussian", nitt=1500000, burnin=50000, thin=500, data=phenotypic_data_leaftraits, pr=TRUE, verbose=F) #the pheno is explain by the year as fixed because few number of parameters and by the Population and Plant factors, we are following a poisson law due to the count type of the response variable (the pheno)
}, mc.cores=1)

#save the chains
leafthickness_model_wo_plantID_wo_sex<- leafthickness
#save(leafthickness_model_wo_plantID_wo_sex,file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/leafthickness_model_wo_plantID_wo_sex.rda")
```

```{r load the MCMC leafthickness}
load("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/leafthickness_model_wo_plantID.rda")
leafthickness_model_wo_plantID <-leafthickness_model_wo_plantID
```

# Evaluation of the models

To evaluate the performance of the models, we investigated several aspects: 

## Autocorrelation within MCMC

First, we checked the autocorrelation between values within the same MCMC chain to ensure that our thinning interval was sufficient.
```{r autocorrelation values MCMC, eval=F, echo=T}
list_name<- c("Population","Genet")
list_model <- c("growth","growth_pheno","repro_pheno","leafthickness")#"growth_bis","LMA","StomDens","D13C"

for(i in 1:length(list_model)){
  
  model_pheno <- get(paste0(list_model[i],"_model_wo_plantID"))
  name_pheno <- list_model[i]
for(x in 1:2){
  name <- list_name[x]
  
  acf_list <- lapply(model_pheno, function(model) {
  # Extract the MCMC samples for the random effects of provenance
  mcmc_samples <- mcmc(model$VCV[, name])  
  # Calculate ACF
  acf_values <- acf(as.numeric(mcmc_samples), plot = FALSE)  # Convert to numeric if needed
  return(acf_values)
})
  
  # Plot the first ACF
plot(acf_list[[1]], main=paste0("Autocorrelation for ",name," Random Effect ",name_pheno), col="black", lwd=2)

colors <- c("red", "blue", "green")  # Colors for the additional chains
for (i in 2:length(acf_list)) {
  lines(acf_list[[i]]$lag, acf_list[[i]]$acf, col=colors[i-1], lwd=2)
}
# Add a legend to distinguish between the chains
legend("topright", legend=paste("Chain", 1:length(model_pheno)), col=c("black", colors), lwd=2)

plot
#save
png(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/figures_wo_plantID/checkmodel/autocorrelation/Autocorrelation_period_",name,"_",name_pheno,".png"));plot(acf_list[[1]], main=paste0("Autocorrelation for ",name," Random Effect ",name_pheno), col="black", lwd=2)
colors <- c("red", "blue", "green")
for (i in 2:length(acf_list)) {
  lines(acf_list[[i]]$lag, acf_list[[i]]$acf, col=colors[i-1], lwd=2)
}
legend("topright", legend=paste("Chain", 1:length(model_pheno)), col=c("black", colors), lwd=2);dev.off()
  }
}
```


```{r Rmardown 1, include=FALSE}
list_name<- c("Population","Genet")
list_model_rmarkdown <- c("growth")
for(i in 1:length(list_model_rmarkdown)){
  model_pheno <- get(paste0(list_model_rmarkdown[i],"_model_wo_plantID"))
  name_pheno <- list_model_rmarkdown[i]
for(x in 1:2){
  name <- list_name[x]
  acf_list <- lapply(model_pheno, function(model) {
  # Extract the MCMC samples for the random effects of provenance
  mcmc_samples <- mcmc(model$VCV[, name])  
  # Calculate ACF
  acf_values <- acf(as.numeric(mcmc_samples), plot = FALSE)  # Convert to numeric if needed
  return(acf_values)
})
  # Plot the first ACF
plot(acf_list[[1]], main=paste0("Autocorrelation for ",name," Random Effect ",name_pheno), col="black", lwd=2)

colors <- c("red", "blue", "green")  # Colors for the additional chains
for (i in 2:length(acf_list)) {
  lines(acf_list[[i]]$lag, acf_list[[i]]$acf, col=colors[i-1], lwd=2)
}
# Add a legend to distinguish between the chains
legend("topright", legend=paste("Chain", 1:length(model_pheno)), col=c("black", colors), lwd=2)
plot
  }
}
```
  
## Chains convergence  
  
Then, we ensured that the MCMC chains were converging.
```{r plot of MCMC chains, echo=T, eval=F}
for(i in 1:length(list_model)){
model_pheno <- get(paste0(list_model[i],"_model_wo_plantID"))
  name_pheno <- list_model[i]

for(x in 1: length(list_name)){
  
  name <- list_name[x]
  
  # Create MCMC lists for overall random effects (aggregated)
mcmc_list <- mcmc.list(lapply(model_pheno, function(model) {
  # Aggregate random effects for provenance
  mcmc(model$VCV[, name])
}))

par(mfrow=c(2,1), mar=c(4,4,2,1))  # Two plots: one for provenance, one for plant ID

# Plot trace plot for overall provenance random effect
plot(mcmc_list, auto.layout=FALSE, main=paste0("Trace Plot for Overall ",name," Random Effect ",name_pheno), col=c("red", "blue", "green", "black"))

#save
png(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/figures_wo_plantID/checkmodel/convergence/MCMC_period_",name,"_",name_pheno,".png"));par(mfrow=c(2,1), mar=c(4,4,2,1))
plot(mcmc_list, auto.layout=FALSE, main=paste0("Trace Plot for Overall ",name," Random Effect ",name_pheno), col=c("red", "blue", "green", "black"));dev.off()
  }
}
```

```{r Rmarkdown 2, include=FALSE}
for(i in 1:length(list_model_rmarkdown)){
model_pheno <- get(paste0(list_model_rmarkdown[i],"_model_wo_plantID"))
  name_pheno <- list_model_rmarkdown[i]
for(x in 1: length(list_name)){
  name <- list_name[x]
  # Create MCMC lists for overall random effects (aggregated)
mcmc_list <- mcmc.list(lapply(model_pheno, function(model) {
  # Aggregate random effects for provenance
  mcmc(model$VCV[, name])
}))
par(mfrow=c(2,1), mar=c(4,4,2,1))  # Two plots: one for provenance, one for plant ID
# Plot trace plot for overall provenance random effect
plot(mcmc_list, auto.layout=FALSE, main=paste0("Trace Plot for Overall ",name," Random Effect ",name_pheno), col=c("red", "blue", "green", "black"))
  }
}
```

Graphically, the chains appear to converge. We calculated the Gelman-Rubin index to further investigate this: 
```{r evaluating the models}
list_model <- c("growth","growth_pheno","repro_pheno","leafthickness")#"growth_bis","LMA","StomDens","D13C"

gelman_rubin_results <- list()

for(i in 1:length(list_model)){
  
  # Get the model object
  model_pheno <- get(paste0(list_model[i], "_model_wo_plantID"))
  name_pheno <- list_model[i]
  
  # Combine the MCMC chains into a single mcmc.list
  chaine_mc <- lapply(model_pheno, function(m) m$Sol)
  chaine_mc <- do.call(mcmc.list, chaine_mc)
  
  # Gelman-Rubin (GR) criterion to assess convergence across MCMC chains
  GRcriterion_index <- gelman.diag(chaine_mc)
  
  # Store the summary of the GR criterion in the list
  gelman_rubin_results[[name_pheno]] <- list(
    mpsrf = GRcriterion_index$mpsrf
  )
}
gelman_rubin_results
```

We can see that the Gelman-Rubin values are below 1.1, which is defined as the threshold for convergence.

## Similarity of the posterior distribution across chains

We also looked at the posterior distribution to investigate the convergence of the MCMC results.
```{r posterior distribution, echo=T, eval=F}
for(i in 1:length(list_model)){
  model_pheno <- get(paste0(list_model[i],"_model_wo_plantID"))
  name_pheno <- list_model[i]

for(x in 1:length(list_name)){
  
  name <-list_name[x]
  
  density_list <- lapply(model_pheno, function(model) {
  # Extract the MCMC samples for the random effects of provenance
  mcmc_samples <- as.numeric(mcmc(model$VCV[, name])) 
  # Calculate density
  density_values <- density(mcmc_samples)
  return(density_values)
})
  #limit of the plot
  max_density_value <- max(sapply(density_list, function(d) max(d$y)))
  
  # Plot the first density
plot(density_list[[1]], main=paste0("Posterior distribution of the ",name," Random Effect value ",name_pheno), col="black", lwd=2, xlab="Value", ylab="Density",ylim=c(0, max_density_value * 1.02))

# Overlay the remaining densities
colors <- c("red", "blue", "green")  # Colors for the additional chains
for (i in 2:length(density_list)) {
  lines(density_list[[i]], col=colors[i-1], lwd=2)
}

# Add a legend to distinguish between the chains
legend("topright", legend=paste("Chain", 1:length(model_pheno)), col=c("black", colors), lwd=2)

#save
png(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/figures_wo_plantID/postdistrib/Posterior_distrib_period_",name,"_",name_pheno,".png"));plot(density_list[[1]], main=paste0("Posterior distribution of the ",name," Random Effect value ",name_pheno), col="black", lwd=2, xlab="Value", ylab="Density",ylim=c(0, max_density_value * 1.02))
colors <- c("red", "blue", "green")  # Colors for the additional chains
for (i in 2:length(density_list)) {
  lines(density_list[[i]], col=colors[i-1], lwd=2)
}
legend("topright", legend=paste("Chain", 1:length(model_pheno)), col=c("black", colors), lwd=2);dev.off()
  }
}
```

```{r Rmarkdown 3, include=FALSE}

for(i in 1:length(list_model_rmarkdown)){
  model_pheno <- get(paste0(list_model_rmarkdown[i],"_model_wo_plantID"))
  name_pheno <- list_model_rmarkdown[i]
for(x in 1:length(list_name)){
  name <-list_name[x]
  density_list <- lapply(model_pheno, function(model) {
  # Extract the MCMC samples for the random effects of provenance
  mcmc_samples <- as.numeric(mcmc(model$VCV[, name]))
  # Calculate density
  density_values <- density(mcmc_samples)
  return(density_values)
})
  #limit of the plot
  max_density_value <- max(sapply(density_list, function(d) max(d$y)))
  # Plot the first density
plot(density_list[[1]], main=paste0("Posterior distribution of the ",name," Random Effect value ",name_pheno), col="black", lwd=2, xlab="Value", ylab="Density",ylim=c(0, max_density_value * 1.02))

# Overlay the remaining densities
colors <- c("red", "blue", "green")  # Colors for the additional chains
for (i in 2:length(density_list)) {
  lines(density_list[[i]], col=colors[i-1], lwd=2)
}
# Add a legend to distinguish between the chains
legend("topright", legend=paste("Chain", 1:length(model_pheno)), col=c("black", colors), lwd=2)
  }
}
```

# Blups values

Finally, we extracted the BLUPs values from the models: 

To calculate the BLUPs, we used the mean BLUP values across the four MCMC runs.
```{r blups estimate}

#list_model <- c("growth","growth_bis","growth_pheno","repro_pheno","leafthickness","LMA","StomDens","D13C")
list_model <- c("growth","growth_pheno","repro_pheno","leafthickness")#"growth_bis","LMA","StomDens"

for(i in 1:length(list_model)){
  
  model_pheno <- get(paste0(list_model[i],"_model_wo_plantID"))
  name_pheno <- list_model[i]
  
#BLUPs values
for(x in 1:4){
  
  model <- model_pheno[[x]] #extract the chains from the model 
  blups<-data.frame(posterior.mode(model$Sol))
  assign(paste0("blups_",name_pheno,"_",x),blups)
  }
}
  
for(i in 1:length(list_model)){
  
  name_pheno <- list_model[i]
  
  MC1 <- get(paste0("blups_",name_pheno,"_1"))
  MC2 <- get(paste0("blups_",name_pheno,"_2"))
  MC3 <- get(paste0("blups_",name_pheno,"_3"))
  MC4 <- get(paste0("blups_",name_pheno,"_4"))

  Blups_df <- data.frame(MC1,MC2,MC3,MC4)
  Blups_df$mean_blup <- rowMeans(Blups_df)
  
#write_xlsx(data.frame(row.names(Blups_df),Blups_df[,c(5)]),paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/BLUPs_mean/Blups_",name_pheno,"final_period.xlsx"))
}
```


# Draft: 

Traits: 

Basal area

```{r Basal area model, eval=FALSE, include=FALSE}
set.seed(3)

#set the prior info for the fixed variables: year and the randoms ones: Population and Plant
prior_Basal_area=list(R=list(V=1, nu=0.002),G=list((G1=list(V=1, nu=0.002)),(G2=list(V=1, nu=0.002)),(G3=list(V=1, nu=0.002))))

#we can perform the model
Basal_area<- mclapply(1:4, function(i) {#we can perform 4 MCMC 
  Basal_area<-MCMCglmm(SumAB.2021.cm2~1,random=~Population + Genet +StemLength.cm,prior=prior_Basal_area, family="gaussian", nitt=1500000, burnin=50000, thin=500, data=phenotypic_data_growth_bis, pr=TRUE, verbose=F) #the pheno is explain by the year as fixed because few number of parameters and by the Population and Plant factors, we are following a gaussian law due to the quantitative continuous type of the response variable (the pheno)
}, mc.cores=1)

#save the chains
Basal_area_model_wo_plantID<- Basal_area
#save(Basal_area_model_wo_plantID,file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/Basal_area_model_wo_plantID.rda")
```

```{r load the MCMC growth_bis, eval=FALSE, include=FALSE}
load("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/Basal_area_model_wo_plantID.rda")
growth_bis_model_wo_plantID <- Basal_area_model_wo_plantID
```

LMA

```{r LMA model, eval=FALSE, include=FALSE}
set.seed(3)
#set the prior info for the fixed variables: year and the randoms ones: Population and Plant
prior_leaftraits=list(R=list(V=1, nu=0.002),G=list((G1=list(V=1, nu=0.002)),(G2=list(V=1, nu=0.002))))

phenotypic_data_leaftraits$Sex<-as.factor(phenotypic_data_leaftraits$Sex)
#we can perform the model
LMA<- mclapply(1:4, function(i) {#we can perform 4 MCMC 
  LMA<-MCMCglmm(mean_LMA~1,random=~Population + Genet,prior=prior_leaftraits, family="gaussian", nitt=1500000, burnin=50000, thin=500, data=phenotypic_data_leaftraits, pr=TRUE, verbose=F) #the pheno is explain by the year as fixed because few number of parameters and by the Population and Plant factors, we are following a poisson law due to the count type of the response variable (the pheno)
}, mc.cores=1)

#save the chains
LMA_model_wo_plantID_wo_sex<- LMA
#save(LMA_model_wo_plantID_wo_sex,file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/LMA_model_wo_plantID_wo_sex.rda")
```

```{r load the MCMC LMA, eval=FALSE, include=FALSE}
load("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/LMA_model_wo_plantID.rda")

LMA_model_wo_plantID <-LMA_model_wo_plantID
```


StomDens

```{r StomDens model, eval=FALSE, include=FALSE}
set.seed(3)
#set the prior info for the fixed variables: year and the randoms ones: Population and Plant
prior_leaftraits=list(R=list(V=1, nu=0.002),G=list((G1=list(V=1, nu=0.002)),(G2=list(V=1, nu=0.002))))

phenotypic_data_leaftraits$Sex<-as.factor(phenotypic_data_leaftraits$Sex)
#we can perform the model
StomDens<- mclapply(1:4, function(i) {#we can perform 4 MCMC 
  StomDens<-MCMCglmm(mean_StomDens~1,random=~Population + Genet,prior=prior_leaftraits, family="gaussian", nitt=1500000, burnin=50000, thin=500, data=phenotypic_data_leaftraits, pr=TRUE, verbose=F) #the pheno is explain by the year as fixed because few number of parameters and by the Population and Plant factors, we are following a poisson law due to the count type of the response variable (the pheno)
}, mc.cores=1)

#save the chains
StomDens_model_wo_plantID_wo_sex<- StomDens
#save(StomDens_model_wo_plantID_wo_sex,file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/StomDens_model_wo_plantID_wo_sex.rda")
```

```{r load the MCMC Stomdens, eval=FALSE, include=FALSE}
load("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/StomDens_model_wo_plantID_wo_sex.rda")

StomDens_model_wo_plantID <- StomDens_model_wo_plantID_wo_sex
```


D13C

```{r D13C model,eval=FALSE, include=FALSE}
set.seed(3)
#set the prior info for the fixed variables: year and the randoms ones: Population and Plant
prior_leaftraits=list(R=list(V=1, nu=0.002),G=list((G1=list(V=1, nu=0.002)),(G2=list(V=1, nu=0.002))))

phenotypic_data_leaftraits$Sex<-as.factor(phenotypic_data_leaftraits$Sex)
#we can perform the model
D13C<- mclapply(1:4, function(i) {#we can perform 4 MCMC 
  D13C<-MCMCglmm(mean_D13C~Sex,random=~Population + Genet,prior=prior_leaftraits, family="gaussian", nitt=1500000, burnin=50000, thin=500, data=phenotypic_data_leaftraits, pr=TRUE, verbose=F) #the pheno is explain by the year as fixed because few number of parameters and by the Population and Plant factors, we are following a poisson law due to the count type of the response variable (the pheno)
}, mc.cores=1)

#save the chains
D13C_model_wo_plantID<- D13C
#save(D13C_model_wo_plantID,file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/D13C_model_wo_plantID.rda")
```

```{r load the MCMC,eval=FALSE, include=FALSE}
load("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/mean_years/D13C_model_wo_plantID.rda")
```