---
title: "Redundancy_analyses_candidate_detection"
author: "Thomas Francisco"
date: "2024-04-30"
output:
  html_document:
    number_sections: true #titles
    toc: true #table of content
    toc_float: true # enable the toc to be on the side of the text, always visible
    collapsed: True #control if the toc label will only display top level titles
    toc_depth: 3
---

```{r setup, include=F}
rm(list = ls())
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache = FALSE
)
library(vegan)
library(dplyr)
library(robust) #for rdadapt
library(qvalue) #for rdadapt
library(ggplot2)
library(radiant.data) #for row_names_to_columns
library(textshape) #for colnames_to_row.names
library(writexl)
```

# Introduction

**Candidate selection: Redundancy analysis(RDA)**

In this script we performed outlier detection using Redundancy Analysis (RDA), which does not take population structure into account, and Partial Redundancy Analysis (pRDA), which does.
RDA is a multivariate canonical analysis between principal component analysis (PCA) and multiple regression. It performs analyses similar to multiple regression, but RDA deals with a matrix of response variables rather than just one. It is considered canonical because it produces canonical axes similar to those found in PCA.
The process starts with linear regressions between the explanatory variables and the response variables. We then perform PCA on the fitted values (the results of the linear regressions) to create the canonical axes (RDA axes) to reduce the number of dimensions. These canonical axes represent linear combinations of the explanatory variables. In summary, RDA combines aspects of both PCA and multiple regression, resulting in PCA-like outputs based on fitted values that allow us to extrapolate the model according to a linear relationship.  
Thus, RDA is a multivariate analysis following a linear hypothesis.

pRDA is a variant of RDA that corrects for covariates during the analysis. The steps are identical to those in RDA, except that an additional step is performed first. The aim of pRDA is to perform RDA while adjusting for other variables that may influence the response variable and/or the explanatory variables. To achieve this, pRDA first performs linear regressions between the response variables and the conditioned variables (the additional variables that could affect the response/explanatory variables). It also regresses the explanatory variables against the covariates. The same steps as in RDA are then performed, but on the residuals of both the response and the explanatory variables. This approach allows the effects of the conditioned variables on the response and explanatory variables to be removed.

*Sources: Legendre and Legendre 2012; Capblancq 2021; workshop on redundancy analysis by the Quebec Center for Biodiversity Science (link).*

# Data
    
The genomic data are the allele frequencies at the population level (29 pop) on the imputed data set (475 indiv, 8616 SNPs) with MAC correction.
```{r data} 
#genomic data
load("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/genetic_new/data_allelic_frequencies_29pop_adapcon_gentree_475_8616.Rdata")

genomic_matrix <- data_allelic_frequencies_29pop_adapcon_gentree_475_8616  

#climatic/structure/IBD
load("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/GEA_new_var/variance_partitioning/Climatic_data_RDA_pRDA.Rdata") 
```

## Models

We perform RDA using the genomic matrix corrected for MAC and the scaled climatic variables of interest. For pRDA, we also included the first 2 Pcs axes.
```{r RDA without correcting, message=FALSE, warning=FALSE}
#model
RDA_env <- rda(formula = genomic_matrix ~  Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality, data = Climatic_data_RDA_pRDA, scale=F)

RsquareAdj(RDA_env)
#other analysis to test the model
#summary(RDA_env)
#significativity of the model
#anova.cca(RDA_env)
#test the significativity of the axis and the climatic variables using permutation
#anova.cca(RDA_env, step = 1000, by = "axis")
#anova.cca(RDA_env, step = 1000, by = "term")
```

Interpretation: The R² value of the climate variables on the genomic matrix is quite important, suggesting that it may be interesting to perform genome-environment association (GEA) analyses on this dataset, as it suggests that a significant number of candidate loci can be identified.

```{r pRDA with population correction, message=FALSE, warning=FALSE}
#model
pRDA_env <- rda(formula = genomic_matrix ~  Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality + Condition(PC1+PC2), data = Climatic_data_RDA_pRDA, scale=F)

RsquareAdj(pRDA_env)

#other analysis to test the model
#summary(pRDA_env)
#significativity of the model
#anova.cca(pRDA_env)
#test the significativity of the axis and the climatic variables using permutation
#anova.cca(pRDA_env, step = 1000, by = "axis")
#anova.cca(pRDA_env, step = 1000, by = "term")
```

Next, we conducted candidate selection based on these models. To achieve this, we followed the procedure developed by Capblancq and Forester (2021) and identified outliers based on their positions along a distribution of Mahalanobis distances. These distances are estimated between each locus and the centre of the RDA space using a given number of canonical axes (K).

First, we need to determine the number of canonical axes (RDA axes) that we will retain.
```{r screeplot, warning=FALSE}
list_models <- c("RDA","pRDA")

for(i in 1:length(list_models)){
  
  model <- get(paste0(list_models[i],"_env"))

#screeplot  
  
#in hist
explained_variance <- model$CCA$eig * 100 / sum(model$CCA$eig)
  
scree_data <- data.frame(Axis = factor(1:length(explained_variance), 
                                       labels = paste0("RDA", 1:length(explained_variance))),
                         Variance = explained_variance)
# Plot with ggplot2
plot <- ggplot(scree_data, aes(x = Axis, y = Variance)) +
  geom_bar(stat = "identity", fill = "darkgrey") +
  labs(title = paste("Screeplot -", list_models[i]), 
       x = "RDA Axis", y = "Explained Variance (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),panel.background = element_blank())

print(plot)
#explained variance along each RDA axis
model$CCA$eig*100/sum(model$CCA$eig)

pdf(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/GEA_new_var/Redundancy_analyses/screeplot_",list_models[i],".pdf"))
print(plot);dev.off()
}
```

Interpretation: based on these analyses, we will keep the first two RDA and pRDA axes to perform the candidate detection.

An interesting aspect of redundancy analyses is that we can visualise the populations (or individuals or genotypes, depending on the scale of our study), loci and climatic variables in the same space. 
First, we plotted the populations and climatic variables in a biplot within the RDA space.
```{r population representation RDA space}
list_models <- c("RDA","pRDA")

for(i in 1:length(list_models)){
  
   model <- get(paste0(list_models[i],"_env"))
#score along the 2 first RDA axis
score_climatic_var <- as.data.frame(scores(model, choices=c(1:2), display="bp"))
score_pop_var <- data.frame(model$CCA$u[,c(1,2)])

#meta_data
meta_data_pop <- read.csv("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Populations/taxus_sample_29pop.csv",h=T,sep=";",dec=",")

#alphabetic order
meta_data_pop_order <- meta_data_pop[order(meta_data_pop$Population),]

#merge for country info
Score_population_bis <- rownames_to_column(score_pop_var,"Population")
score_with_country_info <- merge(Score_population_bis,meta_data_pop_order[,c(1,2)],"Population")
score_with_country_info$Country <- as.factor(score_with_country_info$Country)

#explained variance along each RDA axis
explained_variance <- data.frame(model$CCA$eig)*100/sum(model$CCA$eig) # variance of each axis*100 divided by the sum of all the explained variance for all axis

explained_variance_round <- round(explained_variance$model.CCA.eig,digits=1)

group_palette <- c("Bosnia"="orangered3", "France"="gold2","Germany"= "darkorchid3", "Greece"="navyblue", "Italy"="turquoise2", "Norway"="green3", "Slovakia"="blue", "Slovenia"="red", "Spain"="black", "Sweden"="gray", "Switzerland"="orange", "UK"="darkgreen")

##Biplot with populations and climatic variables along the 2 first RDA axis
biplot_populations <- ggplot() +
  geom_hline(yintercept = 0, linetype = "dashed", color = gray(0.80), size = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", color = gray(0.80), size = 0.6) +
  geom_point(data = score_with_country_info, aes(x = RDA1 * 3, y = RDA2 * 3, colour = Country), size = 2.5, alpha = 0.8) +
  geom_segment(data = score_climatic_var, aes(xend = RDA1, yend = RDA2, x = 0, y = 0), colour = "black", size = 0.15, linetype = 1, arrow = arrow(length = unit(0.02, "npc"))) +
  geom_text(data = score_climatic_var, aes(x=1.1*RDA1, y=1.1*RDA2, label = row.names(score_climatic_var)), size = 3.5)+
  xlab(paste0("RDA 1 (",explained_variance_round[1],"%)")) + 
  ylab(paste0("RDA 2 (",explained_variance_round[2],"%)")) +
  ggtitle(paste0("Biplot ", list_models[i]," Populations")) +
  scale_color_manual(name = "Countries", values = group_palette, labels = levels(score_with_country_info$Country)) +
   scale_x_continuous(expand = expansion(mult = 0.14)) +  # Expands x-axis space
  scale_y_continuous(expand = expansion(mult = 0.1)) +  # Expands y-axis space
  theme_bw(base_size = 14) +
  theme(legend.position="right", panel.grid = element_blank(), strip.text = element_text(size=14),plot.title = element_text(hjust = 0.5,color = "Black",face="italic"))+
  labs(color = "Country")

 print(biplot_populations)
 pdf(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/GEA_new_var/Redundancy_analyses/biplot_",list_models[i],"_populations.pdf"));print(biplot_populations);dev.off()
}
```

Interpretation:  
In the RDA analysis, we observe that populations located at higher altitudes or in eastern regions are located to the left along the first RDA axis, while populations located at lower altitudes with warmer temperatures are located to the right. This suggests that RDA1 represents a gradient of continentality and altitude.  
For RDA2, the axis appears to separate populations based on their precipitation levels.

pRDA interpretation: 
The axes in the pRDA analysis are more complex to interpret than those in the RDA. There is no clear geographical or altitudinal pattern. The first axis appears to separate populations based on diurnal temperature variation, precipitation seasonality and temperature during the driest quarter. Meanwhile, the second axis appears to separate populations based on temperature seasonality, diurnal temperature range, and annual precipitation and temperature.

# Outlier identification

The next step is to run the genome scan to calculate the Mahalanobis distance. This will allow us to calculate the associated q-values and p-values, which will be used as thresholds for candidate selection.

## Calculating p-values and q-values

We used the rdadapt function from Capblancq and Forester (2021):
```{r genome scan function}
rdadapt<-function(rda,K)
{
  zscores<-rda$CCA$v[,1:as.numeric(K)]
  resscale <- apply(zscores, 2, scale)
  resmaha <- covRob(resscale, distance = TRUE, na.action= na.omit, estim="pairwiseGK")$dist
  lambda <- median(resmaha)/qchisq(0.5,df=K)
  reschi2test <- pchisq(resmaha/lambda,K,lower.tail=FALSE)
  qval <- qvalue(reschi2test)
  q.values_rdadapt<-qval$qvalues
  return(data.frame(p.values=reschi2test, q.values=q.values_rdadapt))
}
```

```{r Pvalues models}

for(i in 1:length(list_models)){
  
   model <- get(paste0(list_models[i],"_env"))

#Perform the function to calculate the mahalanobis distance and then pvalues/qvalues
genome_scan <- rdadapt(model,K=2) #the K is equal to the number of RDA axis that we want to use for the selection

assign(paste0("genome_scan_",list_models[i]), genome_scan)

#save the pvalues for each snp for retaining SNPs in LD with the greater signal in the outliers identification script
pvalues_model_snp <- data.frame(snp_names=colnames(genomic_matrix),pvalues= genome_scan$p.values)

assign(paste0("pvalues_", list_models[i],"_snp"), pvalues_model_snp)

   # Now save the object with a descriptive filename
   save(list = paste0("pvalues_", list_models[i],"_snp"), file = paste0("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/genetic_new/GEA_new_var/pvalues_", list_models[i], "_snp.Rdata"))

load(paste0("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/genetic_new/GEA_new_var/pvalues_",list_models[i],"_snp.Rdata"))

write_xlsx(get(paste0("pvalues_", list_models[i],"_snp")),paste0("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/genetic_new/GEA_new_var/pvalues_",list_models[i],"_snp.xlsx"))

#plot the distribution of pvalues 
Histogram_of_Pvalues_RDA<- hist(genome_scan$p.values,
     main= paste0("Histogram of ",list_models[i]," P-values"),
     xlab= "P-values")

#save the histogram
 png(filename=paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/GEA_new_var/Redundancy_analyses/Histogram_of_Pvalues_",list_models[i],".png"));hist(genome_scan$p.values,
     main= paste0("Histogram of ",list_models[i],"P-values"),
     xlab= "P-values");dev.off()
}
```

Interpretation: The distribution of p values seems to follow the expected distribution characterised by a peak of values around 0 followed by a relatively constant frequency of higher values. 

## Threshold selection
    
Using the calculated p-values and q-values, we can now proceed with candidate selection. This involves setting a threshold to identify significant loci based on their statistical significance.
```{r selection of outliers}
for(i in 1:length(list_models)){
  
   genome_scan_model <- get(paste0("genome_scan_",list_models[i]))
# qvalues < 0.05
thres_qvalues <- 0.05

outliers_qvalues_5perc <- data.frame(Loci = colnames(genomic_matrix)[which(genome_scan_model$q.values<thres_qvalues)], p.value = genome_scan_model$q.values[which(genome_scan_model$q.values<thres_qvalues)], contig = unlist(lapply(strsplit(colnames(genomic_matrix)[which(genome_scan_model$q.values<thres_qvalues)], split = "_"), function(x) x[1])))

nrow(outliers_qvalues_5perc)

assign(paste0("outliers_qvalues_5perc_", list_models[i]), outliers_qvalues_5perc)

# qvalues < 0.10
thres_qvalues <- 0.10

outliers_qvalues_10perc <- data.frame(Loci = colnames(genomic_matrix)[which(genome_scan_model$q.values<thres_qvalues)], p.value = genome_scan_model$q.values[which(genome_scan_model$q.values<thres_qvalues)], contig = unlist(lapply(strsplit(colnames(genomic_matrix)[which(genome_scan_model$q.values<thres_qvalues)], split = "_"), function(x) x[1])))

nrow(outliers_qvalues_10perc)
assign(paste0("outliers_qvalues_10perc_", list_models[i]), outliers_qvalues_10perc)

#threshold pvalues < 0.05 with Bonferonni correction
thres_pvalues <- 0.05/length(genome_scan_model$p.values)

outliers_pvalues <- data.frame(Loci = colnames(genomic_matrix)[which(genome_scan_model$p.values<thres_pvalues)], p.value = genome_scan_model$p.values[which(genome_scan_model$p.values<thres_pvalues)], contig = unlist(lapply(strsplit(colnames(genomic_matrix)[which(genome_scan_model$p.values<thres_pvalues)], split = "_"), function(x) x[1])))

nrow(outliers_pvalues)
assign(paste0("outliers_pvalues_", list_models[i]), outliers_pvalues)

#top 1%
##add colnames
#genome_scan_model$snp_names <- colnames(genomic_matrix)

#outliers_top1perc <- genome_scan_model %>% 
#  arrange(p.values) %>%
#slice(1:(0.01*nrow(.))) %>%  
#  as.data.frame()

#assign(paste0("outliers_top1perc_", list_models[i]), outliers_top1perc)
}
```

## Graphic representations of the outliers

### FDR 5%
    
We can plot both the candidate loci and the presumed neutral loci within the RDA space, incorporating the climatic variables.
```{r mahalanobis distance FDR 0.05 threshold}
for(i in 1:length(list_models)){

   model <- get(paste0(list_models[i],"_env"))
   outliers_qvalues_5perc <- get(paste0("outliers_qvalues_5perc_", list_models[i]))
   
  
score_loci <- as.data.frame(scores(model, choices=c(1:2), display="species", scaling="none"))
score_loci_outliers <- data.frame(names = row.names(score_loci), score_loci)
score_loci_outliers$FDR5 <- "other SNPs"
score_loci_outliers$FDR5[score_loci_outliers$names%in%outliers_qvalues_5perc$Loci] <- "candidate SNPs FDR 5%"
score_loci_outliers$FDR5 <- factor(score_loci_outliers$FDR5, levels = c("candidate SNPs FDR 5%","other SNPs"))
score_loci_outliers <- score_loci_outliers[order(score_loci_outliers$FDR5),]

score_climatic_var <- as.data.frame(scores(model, choices=c(1:2), display="bp"))

explained_variance <- data.frame(model$CCA$eig)*100/sum(model$CCA$eig) # variance of each axis*100 divided by the sum of all the explained variance for all axis
explained_variance_round <- round(explained_variance$model.CCA.eig,digits=1)

#Biplot with SNPs and climatic variables along the two first RDA axis. 
 biplot_outliers<- ggplot() +
  geom_hline(yintercept=0, linetype="dashed", color = gray(.80), size=0.6) +
  geom_vline(xintercept=0, linetype="dashed", color = gray(.80), size=0.6) +
  geom_point(data = score_loci_outliers, aes(x=RDA1*15, y=RDA2*15,colour=FDR5), size = 1.4) +
  geom_segment(data = score_climatic_var, aes(xend=RDA1, yend=RDA2, x=0, y=0), colour="black", size=0.15, linetype=1, arrow=arrow(length = unit(0.02, "npc"))) +
  geom_text(data = score_climatic_var, aes(x=1.1*RDA1, y=1.1*RDA2, label = row.names(score_climatic_var)), size = 3.5) +
  xlab(paste0("RDA 1 (",explained_variance_round[1],"%)")) + 
  ylab(paste0("RDA 2 (",explained_variance_round[2],"%)")) +
   ggtitle(paste0(list_models[i]," space: candidates FDR 5%")) +
  guides(color=guide_legend(title="SNP type")) +
  scale_color_manual(values = c("#F9A242FF","lightblue")) +
    scale_x_continuous(expand = expansion(mult = 0.18)) +  # Expands x-axis space
  scale_y_continuous(expand = expansion(mult = 0.16)) +  # Expands y-axis space
  theme_bw(base_size = 14) +
  theme(legend.position="right", panel.grid = element_blank(), strip.text = element_text(size=14),plot.title = element_text(hjust = 0.5,color = "Black",face="italic"))
 
 print(biplot_outliers)
 
 pdf(paste0("C:/Users/tfrancisco/Documents/Thèse/Results/species/taxus/GEA_new_var/RDA/biplot_outliers_",list_models[i],"_FDR_5perc.pdf"));print(biplot_outliers);dev.off()
}
```

### FDR 10%

We can visualise the SNPs identified under a FDR threshold of 10%

```{r mahalanobis distance FDR 0.10 threshold}

for(i in 1:length(list_models)){

   model <- get(paste0(list_models[i],"_env"))
   outliers_qvalues_10perc <- get(paste0("outliers_qvalues_10perc_", list_models[i]))

score_loci <- as.data.frame(scores(model, choices=c(1:2), display="species", scaling="none"))
score_loci_outliers <- data.frame(names = row.names(score_loci), score_loci)
score_loci_outliers$FDR5_10 <- "other SNPs"
score_loci_outliers$FDR5_10[score_loci_outliers$names%in%outliers_qvalues_10perc$Loci] <- "candidate SNPs FDR 10%"
score_loci_outliers$FDR5_10 <- factor(score_loci_outliers$FDR5_10, levels = c("candidate SNPs FDR 10%","other SNPs"))
score_loci_outliers <- score_loci_outliers[order(score_loci_outliers$FDR5_10),]

score_climatic_var <- as.data.frame(scores(model, choices=c(1:2), display="bp"))

explained_variance <- data.frame(model$CCA$eig)*100/sum(model$CCA$eig) # variance of each axis*100 divided by the sum of all the explained variance for all axis
explained_variance_round <- round(explained_variance$model.CCA.eig,digits=1)

#Biplot with SNPs and climatic variables along the two first RDA axis including the less conservative threshold
 biplot_outliers_RDA_LC<- ggplot() +
  geom_hline(yintercept=0, linetype="dashed", color = gray(.80), size=0.6) +
  geom_vline(xintercept=0, linetype="dashed", color = gray(.80), size=0.6) +
  geom_point(data = score_loci_outliers, aes(x=RDA1*15, y=RDA2*15,colour=FDR5_10), size = 1.4) +
  geom_segment(data = score_climatic_var, aes(xend=RDA1, yend=RDA2, x=0, y=0), colour="black", size=0.15, linetype=1, arrow=arrow(length = unit(0.02, "npc"))) +
  geom_text(data = score_climatic_var, aes(x=1.1*RDA1, y=1.1*RDA2, label = row.names(score_climatic_var)), size = 3.5) +
  xlab(paste0("RDA 1 (",explained_variance_round[1],"%)")) + 
  ylab(paste0("RDA 2 (",explained_variance_round[2],"%)")) +
   ggtitle(paste0(list_models[i]," space: candidates FDR 10%")) +
  guides(color=guide_legend(title="SNP type")) +
  scale_color_manual(values = c("orange","lightblue")) +
  scale_x_continuous(expand = expansion(mult = 0.18)) +  # Expands x-axis space
  scale_y_continuous(expand = expansion(mult = 0.16)) +  # Expands y-axis space
  theme_bw(base_size = 14) +
  theme(legend.position="right", panel.grid = element_blank(), strip.text = element_text(size=14),plot.title = element_text(hjust = 0.5,color = "Black",face="italic"))
 
 print(biplot_outliers_RDA_LC)
 
  pdf(paste0("C:/Users/tfrancisco/Documents/Thèse/Results/species/taxus/GEA_new_var/RDA/biplot_outliers_",list_models[i],"_FDR_10perc.pdf"));print(biplot_outliers);dev.off()
}
```

We can also visualise the top candidate loci, selected based on a false discovery rate (FDR) of 5%, alongside the presumably neutral loci within the RDA space. 
```{r mahalanobis distance FDR 0.10 +0.05 thresholds}

for(i in 1:length(list_models)){

   model <- get(paste0(list_models[i],"_env"))
   outliers_qvalues_10perc <- get(paste0("outliers_qvalues_10perc_", list_models[i]))
   outliers_qvalues_5perc <- get(paste0("outliers_qvalues_5perc_", list_models[i]))

score_loci <- as.data.frame(scores(model, choices=c(1:2), display="species", scaling="none"))
score_loci_outliers <- data.frame(names = row.names(score_loci), score_loci)
score_loci_outliers$FDR5_10 <- "other SNPs"
score_loci_outliers$FDR5_10[score_loci_outliers$names%in%outliers_qvalues_10perc$Loci] <- "candidate SNPs FDR 10%"
score_loci_outliers$FDR5_10[score_loci_outliers$names%in%outliers_qvalues_5perc$Loci] <- "candidate SNPs FDR 5%"
score_loci_outliers$FDR5_10 <- factor(score_loci_outliers$FDR5_10, levels = c("candidate SNPs FDR 10%","candidate SNPs FDR 5%","other SNPs"))
score_loci_outliers <- score_loci_outliers[order(score_loci_outliers$FDR5_10),]

score_climatic_var <- as.data.frame(scores(model, choices=c(1:2), display="bp"))

explained_variance <- data.frame(model$CCA$eig)*100/sum(model$CCA$eig) # variance of each axis*100 divided by the sum of all the explained variance for all axis
explained_variance_round <- round(explained_variance$model.CCA.eig,digits=1)

#Biplot with SNPs and climatic variables along the two first RDA axis including the less conservative threshold
 biplot_outliers_RDA_LC<- ggplot() +
  geom_hline(yintercept=0, linetype="dashed", color = gray(.80), size=0.6) +
  geom_vline(xintercept=0, linetype="dashed", color = gray(.80), size=0.6) +
  geom_point(data = score_loci_outliers, aes(x=RDA1*15, y=RDA2*15,colour=FDR5_10), size = 1.4) +
  geom_segment(data = score_climatic_var, aes(xend=RDA1, yend=RDA2, x=0, y=0), colour="black", size=0.15, linetype=1, arrow=arrow(length = unit(0.02, "npc"))) +
  geom_text(data = score_climatic_var, aes(x=1.1*RDA1, y=1.1*RDA2, label = row.names(score_climatic_var)), size = 3.5) +
  xlab(paste0("RDA 1 (",explained_variance_round[1],"%)")) + 
  ylab(paste0("RDA 2 (",explained_variance_round[2],"%)")) +
   ggtitle(paste0(list_models[i]," space: candidates FDR 10%")) +
  guides(color=guide_legend(title="SNP type")) +
  scale_color_manual(values = c("darkgreen","orange","lightblue")) +
    scale_x_continuous(expand = expansion(mult = 0.18)) +  # Expands x-axis space
  scale_y_continuous(expand = expansion(mult = 0.16)) +  # Expands y-axis space
  theme_bw(base_size = 14) +
  theme(legend.position="right", panel.grid = element_blank(), strip.text = element_text(size=14),plot.title = element_text(hjust = 0.5,color = "Black",face="italic"))
 
 print(biplot_outliers_RDA_LC)
 
  pdf(paste0("C:/Users/tfrancisco/Documents/Thèse/Results/species/taxus/GEA_new_var/RDA/biplot_outliers_",list_models[i],"_FDR_10perc+5perc.pdf"));print(biplot_outliers);dev.off()
}
```
## Save candidates

We can save the candidates: 

### FDR 5% candidates 

```{r save FDR 5% candidates}
for(i in 1:length(list_models)){

  outliers_FDR5perc_T_Adapcon_gentree <- get(paste0("outliers_qvalues_5perc_", list_models[i]))[,-3]
  
  assign(paste0("outliers_",list_models[i],"_FDR5perc_T_Adapcon_gentree"),outliers_FDR5perc_T_Adapcon_gentree)
  
  # Now save the object with a descriptive filename
   save(list = paste0("outliers_",list_models[i],"_FDR5perc_T_Adapcon_gentree"), file = paste0("C:/Users/tfrancisco/Documents/Thèse/Results/species/taxus/GEA_new_var/outliers/outliers_",list_models[i],"FDR5perc_T_Adapcon_gentree.Rdata"))

write_xlsx(get(paste0("outliers_",list_models[i],"_FDR5perc_T_Adapcon_gentree")),paste0("C:/Users/tfrancisco/Documents/Thèse/Results/species/taxus/GEA/outliers/outliers_",list_models[i],"_FDR5perc_T_Adapcon_gentree.xlsx"))
}
```

### FDR 10% candidates 
```{r save FDR 10% candidates}

for(i in 1:length(list_models)){

  outliers_FDR10perc_T_Adapcon_gentree <- get(paste0("outliers_qvalues_10perc_", list_models[i]))[,-3]
  
  assign(paste0("outliers_",list_models[i],"_FDR10perc_T_Adapcon_gentree"),outliers_FDR10perc_T_Adapcon_gentree)
  
  # Now save the object with a descriptive filename
   save(list = paste0("outliers_",list_models[i],"_FDR10perc_T_Adapcon_gentree"), file = paste0("C:/Users/tfrancisco/Documents/Thèse/Results/species/taxus/GEA_new_var/outliers/outliers_",list_models[i],"FDR10perc_T_Adapcon_gentree.Rdata"))

write_xlsx(get(paste0("outliers_",list_models[i],"_FDR10perc_T_Adapcon_gentree")),paste0("C:/Users/tfrancisco/Documents/Thèse/Results/species/taxus/GEA/outliers/outliers_",list_models[i],"_FDR10perc_T_Adapcon_gentree.xlsx"))
}
```


**What follows is a draft detailing another method not used in our study for identifying outliers based on the extremeness of loadings along the RDA axes , as described by Forester et al. (2018).**

We also performed outlier detection based on the loadings of loci along the RDA axes to identify extreme loadings for each retained axis, following the methodology outlined by Forester et al. (2018).

```{r extremness along RDA axis, eval=FALSE, include=FALSE}
load.rda <- scores(RDA_env, choices=c(1:3), display="species")

hist(load.rda[,1], main="Loadings on RDA1")
hist(load.rda[,2], main="Loadings on RDA2")
```

Function from Forester et al. (2018).
```{r perform the detection, eval=FALSE, include=FALSE}
#function to do it
outliers_function <- function(x,z){
  lims <- mean(x) + c(-1, 1) * z * sd(x)     # find loadings +/-z sd from mean loading     
  x[x < lims[1] | x > lims[2]]               # locus names in these tails
}
cand1 <- data.frame(outliers_function(load.rda[,1],3)) # 3
colnames(cand1) <- c("pvalues"); cand1$snp_names <- row.names(cand1)
cand2 <- data.frame(outliers_function(load.rda[,2],3)) # 51
colnames(cand2) <- c("pvalues"); cand2$snp_names <- row.names(cand2)

#names of the outliers
outliers_extremness <- rbind (cand1,cand2)

nrow(outliers_extremness)
```
