---
title: "Relationship_fitness_genomic_offset"
author: "Thomas Francisco"
date: "2024-09-18"
output:
  html_document:
    number_sections: true #titles
    toc: true #table of content
    toc_float: true # enable the toc to be on the side of the text, always visible
    collapsed: True #control if the toc label will only display top level titles
    toc_depth: 3
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	cache = FALSE
)
rm(list = ls())

library(dplyr)
library(ggplot2)
library(ggcorrplot)
library(brms)
library(boot)
library(corrplot)
library(gridExtra)
library(stringr)
library(purrr)
```

# Introduction

To evaluate the relevance of the genomic offset in providing insight into potential maladaptation to climate across studied Taxus baccata populations, we investigated the potential relationship between fitness proxies and genomic offset predictions using an independent dataset of populations not used to build the gene-climate models. Genomic offset values for each population were calculated as the Euclidean distance between the predicted optimal genomic compositions under the population's reference climate and the clonal bank's climate from the planting to the measurement dates. We used two methods to calculate the genomic offset:  
- RDA method with 5 sets of SNPs  
- Gradient forest with 5 sets of SNPs  
      
For the phenotypic data, we aimed to retain the most fitness-related traits. We selected four traits:
- Growth: Shoot volume
- Growth phenology: Spring elongation
- Reproductive phenology: proportion of male open strobili
- Ecophysiological: leaf thickness

To search for an association between the phenotypic data and the genomic offset predictions, we used several analyses: 
- Pearson correlation
- linear model: inferential and Bayesian approaches
- non linear model: inferential and Bayesian approaches
      
# Data

```{r meta_data}
meta_data <- read.csv("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Populations/Populations_common_garden.csv",h=T,sep=";",dec=",")
```

## BLUPs data 
    
First, we loaded the BLUPs phenotypic data at the population level. We only kept the BLUPs at the population level.   
For growth, growth and reproductive phenology:
```{r phenotypic data}
#list_traits <-c("growth","growth_pheno","repro_pheno","leafthickness","LMA","StomDens","D13C")
list_traits <-c("growth","growth_pheno","repro_pheno")

#selection of the right coeff
list_x <- c(2,2,2)
list_y <- c(27,27,19)

for(i in 1:length(list_traits)){
  
  trait <- list_traits[i]
  X <- list_x[i]
  Y <- list_y[i]
  
  data <- read.csv(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/Blups_mean/Blups_",trait,"_period.csv"),h=T,sep=";",dec=",")

  data_pop <- data[X:Y,] #keep only the Blup at the population level 
  colnames(data_pop) <- c("Population","Blups")
  data_pop$Population <- str_remove(data_pop$Population, "Population\\.")
  
  assign(paste0("df_",trait),data_pop)
}
```

For leaf thickness:
```{r phenotypic leaf traits}
list_traits <-c("leafthickness")#"growth_bis","StomDens","LMA"

#selection of the right coeff
list_x <- c(2,2,2,2)
list_y <- c(27,27,27,27)

for(i in 1:length(list_traits)){
  
  trait <- list_traits[i]
  X <- list_x[i]
  Y <- list_y[i]
  
  data <- read.csv(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/phenotypic_data/Blups_mean/Blups_",trait,"final_period.csv"),h=T,sep=";",dec=",")

  data_pop <- data[X:Y,] #keep only the Blup at the population level 
  colnames(data_pop) <- c("Population","Blups")
  data_pop$Population <- str_remove(data_pop$Population, "Population\\.")
  
  assign(paste0("df_",trait),data_pop)
}
```

We removed the populations with too few individuals measured
```{r removing the populations with to few individuals}
#Shootvol
#We remove the ACEBDO, LASENIA pop
df_growth_final <- df_growth %>%
  filter(!(Population %in% c("ACEBEDO", "LASENIA")))

#springelong
#We remove the ACEBEDO, ALCOY, LASENIA and QUESADA pop

df_growth_pheno_final <- df_growth_pheno %>%
  filter(!(Population %in% c("ACEBEDO", "ALCOY", "LASENIA","QUESADA")))

#totalstrobopen
#We remove the BOCA, JERTE, LASENIA, PUBLA and RASQUERA
df_repro_pheno_final <- df_repro_pheno %>%
  filter(!(Population %in% c("BOCAHUERGANO", "JERTE", "LASENIA","	PUEBLALILLO","RASQUERA")))

#leaves traits
#remove ACEBEDO, ALCOY, LASENIA and QUESADA
##leafthickness
df_leafthickness_final <- df_leafthickness %>%
  filter(!(Population %in% c("ACEBEDO", "ALCOY", "LASENIA","	QUESADA")))
```

## Correlation phenotypic traits

We investigated the correlation of the phenotypic traits of populations: 

```{r correlation matrix data}

colnames(df_growth_final) <- c("Population","Growth")
colnames(df_growth_pheno_final) <- c("Population","Growth_pheno")
colnames(df_repro_pheno_final) <- c("Population","Repro_pheno")
colnames(df_leafthickness_final) <- c("Population","Leafthickness")

merged_df <- df_growth_final %>%
  full_join(df_growth_pheno_final, by = "Population") %>%
  full_join(df_repro_pheno_final, by = "Population") %>%
  full_join(df_leafthickness_final, by = "Population")
```

```{r correlation}
correlation_phenotypic_traits <- cor(merged_df[,-1], use="pairwise.complete.obs")

corrplot(correlation_phenotypic_traits, method = "number", addrect = 2, col = c("darkorange","darkred"), type = "lower", tl.col = "black", tl.cex = 0.6, number.cex = 0.6)
```

We can see that the phenotypic traits do not seem to be highly related, as the correlations are quite low. This can indicate that they are responding to different selective pressures? Trade-off? Also, does that mean that they provide complementary information about fitness? 


## Composite fitness index

We also tried to integrate these different traits into a single index because fitness is multitraits especially for long-lived organisms such as trees. 
The idea is to organize the populations based on their trait values and what seems to be adaptive in the common garden. For example, high growth is advantageous, so the population with the highest growth among the 26 populations will have the highest value. These values will be scaled to give equal weight to each trait, as we do not have the same number of populations per trait (using rank values would give more weight to traits with more populations).
To standardize the values, we are applying the formula:   
                              Standardized trait= (trait - min_value_trait)/(max_value_trait - min_value_trait)
                              
```{r standardized values}
list_traits <- c("growth","growth_pheno","repro_pheno","leafthickness")

for(x in 1: length(list_traits)){
  
  trait <- list_traits[x]
  df <- get(paste0("df_", trait,"_final"))
  
  df_standardized <- df %>%
  mutate(
    trait_scaled = (df[,2] - min(df[,2])) / (max(df[,2]) - min(df[,2])))
  
  colnames(df_standardized) <- c("Population","Blups",paste0(trait,"_standardized"))
  
  assign(paste0("df_standardized_", trait), df_standardized)
}
```

```{r standardize df}
final_tot_standardized_traits <- reduce(list(df_standardized_growth[,-2], df_standardized_growth_pheno[,-2], df_standardized_repro_pheno[,-2], df_standardized_leafthickness[,-2]), full_join, by = "Population")
```

```{r composite fitness calculation}
Trait_calculation <- final_tot_standardized_traits %>%
  rowwise() %>%
  mutate(
    # Calculate the sum of standardized traits, ignoring NAs
    valid_traits_sum = sum(c_across(ends_with("_standardized")), na.rm = TRUE),
    
    # Count non-NA standardized traits
    valid_traits_count = sum(!is.na(c_across(ends_with("_standardized")))),
    
    # Calculate the trait index only if there is at least one valid trait
    trait_index = if_else(valid_traits_count > 0, 
                          valid_traits_sum / valid_traits_count, 
                          NA_real_)
  ) %>%
  dplyr::select(-valid_traits_sum, -valid_traits_count) %>%  # Optionally remove helper columns
  ungroup()

# View the combined data frame with trait index
print(Trait_calculation)

df_composite_fitness <- data.frame(Population=Trait_calculation[,1],Fitness_index=Trait_calculation$trait_index)
```

## GO predictions
    
We loaded the genomic offset predictions for the 2 genomic offset methods: 
```{r Load GO predictions}
#RDA standard GO
list_RDA_standard <- c("all","random","random_same_AF","LC","CG")
list_period <- c("CG2012","CG2021")
for(i in 1: length(list_period)){
  
    period <- list_period[i]

for(x in 1:length(list_RDA_standard)){
  name <- list_RDA_standard[x]

  
  load(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/standard_GO/RDA/data/CG_period/GO_RDA_",name,"_",period,"_Taxus.Rdata"))
  }
}

#GF GO
list_GF_standard <- c("all","random","random_same_AF","LC","CG")
list_period <- c("2012","2021")
for(i in 1: length(list_period)){
  
   period <- list_period[i]
  
for(x in 1:length(list_GF_standard)){
  
  name <- list_GF_standard[x]

  
  load(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/Standard_GO/GF/data/CG_period/GO_standard_GF_",name,"_",period,".Rdata"))
}
}
```

We can investigate the correlation between the predictions across methods and models: 
```{r correlation GO predictions}
df_GO <- data.frame(RDA_all=GO_RDA_all_CG2012_Taxus$GO_all_CG2012,RDA_ran=GO_RDA_random_CG2012_Taxus$GO_random_CG2012,RDA_rand_AF=GO_RDA_random_same_AF_CG2012_Taxus$GO_random_same_AF_CG2012,RDA_LC=GO_RDA_LC_CG2012_Taxus$GO_LC_CG2012,RDA_CG=GO_RDA_CG_CG2012_Taxus$GO_CG_CG2012,GF_ran=GO_standard_GF_random_2012$genomic_offset_GF_random,GF_ran_AF=GO_standard_GF_random_same_AF_2012$genomic_offset_GF_random_same_AF,GF_LC=GO_standard_GF_LC_2012$genomic_offset_GF_LC,GF_CG=GO_standard_GF_CG_2012$genomic_offset_GF_CG)

df <- cor(df_GO)
ggcorrplot(df, hc.order = F, type = "lower", lab = TRUE, title = "Correlation Matrix between Genomic Offsets across models")
```

```{r save corrplot GO, include=FALSE}
pdf("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models/correlation/correlation_GO_predictions_taxus.pdf");ggcorrplot(df, hc.order = F, type = "lower", lab = TRUE, title = "Correlation Matrix between Genomic Offsets across models");dev.off()
```

# Relationship fitness ~ genomic offset

## Pearson correlations
        
First, we can investigate whether there is a correlation between fitness values and genomic offset (GO) predictions.
```{r Pearson correlation}
# List of RDA genomic offsets for 2012 and 2021
list_GO_2012_RDA <- c("GO_RDA_all_CG2012_Taxus","GO_RDA_random_CG2012_Taxus", "GO_RDA_random_same_AF_CG2012_Taxus", 
                      "GO_RDA_LC_CG2012_Taxus")
list_GO_2021_RDA <- c("GO_RDA_all_CG2021_Taxus","GO_RDA_random_CG2021_Taxus", "GO_RDA_random_same_AF_CG2021_Taxus", 
                      "GO_RDA_LC_CG2021_Taxus")

# List of Gradient Forest genomic offsets for 2012 and 2021
list_GO_2012_GF <- c("GO_standard_GF_all_2012","GO_standard_GF_random_2012", "GO_standard_GF_random_same_AF_2012", 
                     "GO_standard_GF_LC_2012")
list_GO_2021_GF <- c("GO_standard_GF_all_2021","GO_standard_GF_random_2021", "GO_standard_GF_random_same_AF_2021", 
                     "GO_standard_GF_LC_2021")

# Combine lists into one for easier processing
list_GO_2012 <- c(list_GO_2012_RDA, list_GO_2012_GF)
list_GO_2021 <- c(list_GO_2021_RDA, list_GO_2021_GF)

# Corresponding methods and models
list_method <- c(rep("RDA", 4), rep("GF", 4))  # 4 RDA models and 4 GF models

list_model <- c("all","random", "random_same_AF", "LC", "all","random", "random_same_AF", "LC")
list_model <- c("all","random", "random_same_AF", "outlier", "all","random", "random_same_AF", "outlier")

# Traits considered
#list_traits <- c("growth","growth_pheno","repro_pheno","leafthickness","LMA","StomDens","D13C")
list_traits <- c("growth","growth_pheno","repro_pheno","leafthickness","composite_fitness")

# Create an empty dataframe to store correlations
correlation_df <- data.frame(
  Method = character(),
  Model = character(),
  Trait = character(),
  Correlation = numeric(),
  stringsAsFactors = FALSE
)

# Loop over each Genomic Offset for both RDA and GF models
for(i in 1:8) {  # Now we have 8 models (4 RDA + 4 GF)
  GO_pred_2012 <- get(list_GO_2012[i])
  GO_pred_2021 <- get(list_GO_2021[i])
  method <- list_method[i]
  model <- list_model[i]
  
  # Loop over each trait
  for(x in 1:length(list_traits)) {
    trait <- list_traits[x]
    
    # Use CG2021 data for Height, and CG2012 for other traits
    if (trait == "Leafthickness" || trait=="Composite_fitness") {
      df_trait <- get(paste0("df_", trait))
      
      # Merge data frames based on population for CG2021
      merged_df <- merge(df_trait, GO_pred_2021, by = "Population", suffixes = c(".trait", ".GO"))
    } else {
      df_trait <- get(paste0("df_", trait))
      
      # Merge data frames based on population for CG2012
      merged_df <- merge(df_trait, GO_pred_2012, by = "Population", suffixes = c(".trait", ".GO"))
    }
    
    # Calculate the correlation
    correlation <- cor(merged_df[, 2], merged_df[, 3])
    
    # Store the correlation in the dataframe
    correlation_df <- rbind(correlation_df, data.frame(Method = method, Model = model, Trait = trait, Correlation = correlation))
  }
}
# Print the correlation data frame
#print(correlation_df)
```


Corrplot phenotypic traits, genomic offset
```{r corrplot predicted GO}
# Reshape the data for corrplot if necessary
cor_matrix <- reshape(correlation_df, 
                      timevar = "Trait", 
                      idvar = c("Method", "Model"), 
                      direction = "wide")

# Set rownames based on a combination of 'Method' and 'Model'
rownames(cor_matrix) <- paste(cor_matrix$Method, cor_matrix$Model, sep = "_")

# Remove the 'Method' and 'Model' columns since they are now in the rownames
cor_matrix <- cor_matrix[ , !(names(cor_matrix) %in% c("Method", "Model"))]

colnames(cor_matrix) <- list_traits

# create a correlation plot
cor_matrix_fitness_index <- t(cor_matrix)#to have the traits on the abscissa
ggcorrplot(cor_matrix_fitness_index, hc.order = F, type = "full", lab = TRUE, title = "Correlation Matrix between Traits
           and Genomic Offsets")
```

```{r save corrplot associations, include=FALSE}
pdf("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/correlation/correlation_fitness_index.pdf");ggcorrplot(cor_matrix_fitness_index, hc.order = F, type = "full", lab = TRUE, title = "Correlation Matrix between Traits and Genomic Offsets");dev.off()
```


**We can see that correlations are negatives for almost all the methods and models for each trait. This aligns with what is expected to be advantageous in the climate of the common garden. Indeed, higher growth appears to be beneficial, as does late phenology, since the common garden is situated in a mountainous climate that is sensitive to late frost events. Additionally, the annual amount of precipitation is relatively low compared to the climate of the populations, suggesting that drought tolerance traits, such as high leaf thickness, could also be advantageous.**  
**Depending on the traits, the strength of the correlation varies across models, indicating that no single model seems to perform significantly better than the others. Regarding the traits, the composite index appears to be more closely associated with genomic offset, suggesting that this index could serve as a better fitness proxy, assuming that genomic offset is indeed negatively correlated with fitness.** 
        
## Linear models
        
To further investigate, we built linear models to explore and test for relationships between fitness and genomic offset. We employed two types of models: inferential models and Bayesian models, as Bayesian approaches are known to perform better with small sample sizes.

### Inferential model

```{r inferential linear models, echo=T, eval=F}
list_GO_2012_RDA <- c("GO_RDA_all_CG2012_Taxus","GO_RDA_random_CG2012_Taxus", "GO_RDA_random_same_AF_CG2012_Taxus", 
                      "GO_RDA_LC_CG2012_Taxus")
list_GO_2021_RDA <- c("GO_RDA_all_CG2021_Taxus","GO_RDA_random_CG2021_Taxus", "GO_RDA_random_same_AF_CG2021_Taxus", 
                      "GO_RDA_LC_CG2021_Taxus")

# List of Gradient Forest genomic offsets for 2012 and 2021
list_GO_2012_GF <- c("GO_standard_GF_all_2012","GO_standard_GF_random_2012", "GO_standard_GF_random_same_AF_2012", 
                     "GO_standard_GF_LC_2012")
list_GO_2021_GF <- c("GO_standard_GF_all_2021","GO_standard_GF_random_2021", "GO_standard_GF_random_same_AF_2021", 
                     "GO_standard_GF_LC_2021")

# Combine lists into one for easier processing
list_GO_2012 <- c(list_GO_2012_RDA, list_GO_2012_GF)
list_GO_2021 <- c(list_GO_2021_RDA, list_GO_2021_GF)

# Corresponding methods and models
list_method <- c(rep("RDA", 4), rep("GF", 4))  # 4 RDA models and 4 GF models

list_model <- c("all","random", "random_same_AF", "LC", "all","random", "random_same_AF", "LC")

# Traits considered
list_traits <- c("growth","growth_pheno","repro_pheno","leafthickness","composite_fitness")

# Create an empty dataframe to store correlations
lm_inf_df <- data.frame(
  Method = character(),
  Model = character(),
  Trait = character(),
  R2 = numeric(),
  Coeff = numeric(),
  Pvalue = numeric(),
  stringsAsFactors = FALSE
)

# Loop over each Genomic Offset for both RDA and GF models
for(i in 1:8) {  # Now we have 8 models (4 RDA + 4 GF)
  GO_pred_2012 <- get(list_GO_2012[i])
  GO_pred_2021 <- get(list_GO_2021[i])
  method <- list_method[i]
  model <- list_model[i]
  
  # Loop over each trait
  for(x in 1:length(list_traits)) {
    trait <- list_traits[x]
    
    # Use CG2021 data for Height, and CG2012 for other traits
    if (trait == "leafthickness" || trait== "composite_fitness") {
      df_trait <- get(paste0("df_", trait))
      
      # Merge data frames based on population for CG2021
      merged_df <- merge(df_trait, GO_pred_2021, by = "Population", suffixes = c(".trait", ".GO"))
    } else {
      df_trait <- get(paste0("df_", trait))
      
      # Merge data frames based on population for CG2012
      merged_df <- merge(df_trait, GO_pred_2012, by = "Population", suffixes = c(".trait", ".GO"))
    }
    
    #linear model
    lm_model <- lm(merged_df[, 2] ~ merged_df[, 3])
    
    #assumptions: 
    pdf(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Inferential/plot_model/",method,"/",model,"/diagnostic_plots/",method,"_",model,"_",trait,"_diagnostics.pdf"))
    
    # Plot all four diagnostic plots at once in a 2x2 grid
    par(mfrow = c(2, 2))
    plot(lm_model)
    dev.off()
    
    # Reset plot settings
    par(mfrow = c(1, 1))
    
    coeff <- data.frame(coef(lm_model))[2,1]
    result_model <- data.frame(anova(lm_model))
    r_squared <- summary(lm_model)$r.squared
    pvalues <- result_model[1,5]
    
    # Store the correlation in the dataframe
    lm_inf_df <- rbind(lm_inf_df, data.frame(Method=method,Model = model, Trait = trait,R2=round(r_squared,3),Coeff = round(coeff,3),Pvalue=round(pvalues,3)))
    
    # Create the data frame for plotting
    plot_data <- data.frame(
      GO_values = merged_df[, 3],
      Blup = merged_df[, 2],
      fitted = lm_model$fitted.values
    )
    
    # Generate the ggplot
    p <- ggplot(plot_data, aes(x = GO_values, y = Blup)) +
      geom_point(color = "blue") +
      geom_smooth(method = "lm", se = TRUE, level = 0.95, color = "lightgrey", alpha = 0.2) +  # Add confidence interval bands
      geom_line(aes(y = fitted), color = "red") +
      labs(title = paste("Linear Model:", trait, "~", method, model),
           x = paste0("Genomic offset ",method," ",model),
           y = paste0("proxy fitness ",trait)) +
      theme_minimal()
    
    # Print the plot (or use ggsave() to save each plot to a file)
    #print(p)
    
    pdf(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Inferential/plot_model/",method,"/",model,"/plot_lm_",method,"_",model,"_",trait,"_Taxus.pdf"));print(p);dev.off()
  }
}

# Print the correlation data frame
print(lm_inf_df)
```

```{r linear relationship example rmarkdown, echo=FALSE}
merged_df <- merge(df_composite_fitness, GO_standard_GF_random_same_AF_2021, by = "Population", suffixes = c(".trait", ".GO"))

lm_model <- lm(merged_df[, 2] ~ merged_df[, 3])

summary(lm_model)
coeff <- data.frame(coef(lm_model))[2,1]
    result_model <- data.frame(anova(lm_model))
    r_squared <- summary(lm_model)$r.squared
    pvalues <- result_model[1,5]
    

plot_data <- data.frame(
      GO_values = merged_df[, 3],
      Blup = merged_df[, 2],
      fitted = lm_model$fitted.values
    )
    
    # Generate the ggplot
    p <- ggplot(plot_data, aes(x = GO_values, y = Blup)) +
      geom_point(color = "#FF1493") +
      geom_smooth(method = "lm", se = TRUE, level = 0.95, color = "lightgrey", alpha = 0.2) +  # Add confidence interval bands
      geom_line(aes(y = fitted), color = "blue") +
      labs(title = "Composite fitness = 0.6211 - 8.169xGO + E   R²= 0.25, P-value= 0.014",
           x = "Genomic offset ",
           y = "Composite fitness") +
      theme_minimal()+
      theme_bw(base_size = 11) +
  theme(legend.position="right", panel.grid = element_line(), strip.text = element_text(size=11),plot.title = element_text(hjust = 0.5,color = "Black",face="italic"))
    
    plot(p)
    
#pdf("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Inferential/plot_model/Retained_model/totalstrob_RDA_random_AF_linear_plot.pdf");print(p);dev.off()
```

```{r save lm inf model, include=FALSE}
#lm_inf_df_GO_taxus<- lm_inf_df
#save(lm_inf_df_GO_taxus,file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Inferential/table/lm_inf_df_GO_taxus.Rdata")

load("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Inferential/table/lm_inf_df_GO_taxus.Rdata")
#save table
pdf("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Inferential/table/table_bis.pdf", height=16, width=11.5)
grid.table(lm_inf_df_GO_taxus)
ttheme_default(base_size = 12)
dev.off()
```

We observed that for some traits and models, there is a linear negative relationship. For example, for the composite index and the genomic offset derived from the Gradient Forest (GF) model, using the random set with the same allele frequencies, we found a distinct negative relationship. However, the size of our dataset could be a limitation.

### Bayesian model
      
Based on the observations from the inferential linear model, we also applied a Bayesian linear model. Indeed, the Bayesian framework is generally more effective at handling smaller datasets compared to frequentist approaches. 
```{r linear bayesian model, echo=T, eval=F}
list_GO_2012_RDA <- c("GO_RDA_all_CG2012_Taxus","GO_RDA_random_CG2012_Taxus", "GO_RDA_random_same_AF_CG2012_Taxus", 
                      "GO_RDA_LC_CG2012_Taxus")#, "GO_RDA_CG_CG2012_Taxus"
list_GO_2021_RDA <- c("GO_RDA_all_CG2021_Taxus","GO_RDA_random_CG2021_Taxus", "GO_RDA_random_same_AF_CG2021_Taxus", 
                      "GO_RDA_LC_CG2021_Taxus")#, "GO_RDA_CG_CG2021_Taxus"

# List of Gradient Forest genomic offsets for 2012 and 2021
list_GO_2012_GF <- c("GO_standard_GF_all_2012","GO_standard_GF_random_2012", "GO_standard_GF_random_same_AF_2012", 
                     "GO_standard_GF_LC_2012")#, "GO_standard_GF_CG_2012"
list_GO_2021_GF <- c("GO_standard_GF_all_2021","GO_standard_GF_random_2021", "GO_standard_GF_random_same_AF_2021", 
                     "GO_standard_GF_LC_2021")#, "GO_standard_GF_CG_2021"

# Combine lists into one for easier processing
list_GO_2012 <- c(list_GO_2012_RDA, list_GO_2012_GF)
list_GO_2021 <- c(list_GO_2021_RDA, list_GO_2021_GF)

# Corresponding methods and models
list_method <- c(rep("RDA", 4), rep("GF", 4))  # 5 RDA models and 5 GF models

list_model <- c("all","random", "random_same_AF", "LC", "all","random", "random_same_AF", "LC")#"CG","CG"

# Traits considered
list_traits <- c("growth","growth_pheno","repro_pheno","leafthickness","composite_fitness")

# Create an empty dataframe to store correlations

lm_bayesian_df <- data.frame(
  Method = character(),
  Model = character(),
  Trait = character(),
  R2_model =numeric(),
  R2_CI=numeric(),
  b = numeric(),
  b_CI = numeric(),
  stringsAsFactors = FALSE
)

# Loop over each Genomic Offset for both RDA and GF models
for(i in 1:8) {  # Now we have 8 models (4 RDA + 4 GF)
  GO_pred_2012 <- get(list_GO_2012[i])
  GO_pred_2021 <- get(list_GO_2021[i])
  method <- list_method[i]
  model <- list_model[i]
  
  # Loop over each trait
  for(x in 1:length(list_traits)) {
    trait <- list_traits[x]
    
    # Use CG2021 data for Height, and CG2012 for other traits
    if (trait == "leafthickness" || trait== "composite_fitness" ) {
      df_trait <- get(paste0("df_", trait))
      
      # Merge data frames based on population for CG2021
      merged_df <- merge(df_trait, GO_pred_2021, by = "Population", suffixes = c(".trait", ".GO"))
    } else {
      df_trait <- get(paste0("df_", trait))
      
      # Merge data frames based on population for CG2012
      merged_df <- merge(df_trait, GO_pred_2012, by = "Population", suffixes = c(".trait", ".GO"))
    }

  y_data <- merged_df[,2]
  x_data <- merged_df[,3]
   
df_nl_model <- data.frame(y=y_data,x=x_data)
prior1 <- prior(normal(0, 10), "b1")

fit1 <- brm(
  formula = bf(y ~ x),                  # Formula for the model
  family = gaussian(),                        # Specify Gaussian family
  data = df_nl_model,                         # Data
  prior = prior(normal(0, 10), class = "b",coef = "x"),  # Prior on coefficients (b)
  iter = 1500000,                                # Number of iterations
  warmup = 50000,                                # Warmup iterations
  thin = 500,                                   # Thinning
  cores = 16                                  # Number of cores for parallel computation
)

#save the model: 
assign(paste0("lm_bayesian_",method,"_",model,"_",trait),fit1)
save(list = paste0("lm_bayesian_",method,"_",model,"_",trait),file=paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Bayesian/models/",method,"/lm_bayesian_",method,"_",model,"_",trait,".Rdata"))

#load(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models/Linear_models/Bayesian/models/",method,"/lm_bayesian_",method,"_",model,"_",trait,".Rdata"))

fit1 <- get(paste0("lm_bayesian_",method,"_",model,"_",trait))
#coeff
fit1_summary <- summary(fit1)

# Extract the estimates for b1 and b2
coef_b <- fit1_summary$fixed["x", "Estimate"]

b_ci_lower <- fit1_summary$fixed["x", "l-95% CI"]
b_ci_upper <- fit1_summary$fixed["x", "u-95% CI"]

R2_model <- bayes_R2(fit1)["R2","Estimate"]

R2_ci_lower <- bayes_R2(fit1)["R2", "Q2.5"]
R2_ci_upper <- bayes_R2(fit1)["R2", "Q97.5"]

# Store the correlation in the dataframe
    lm_bayesian_df <- rbind(lm_bayesian_df, data.frame(Method=method,Model = model, Trait = trait, R2_model = round(R2_model,3),R2_CI=paste0(round(R2_ci_lower,3),"_",round(R2_ci_upper,3)),b = round(coef_b,3), b_CI = paste0(round(b_ci_lower,3),"_",round(b_ci_upper,3))))

#MCMC chains
plot(fit1)

pdf(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Bayesian/figures/",method,"/",model,"/eval_model",trait,"_GO_",method,"_",model,"_Taxus.pdf"));plot(fit1);dev.off()

#posterior distrib
pp_check(fit1)

pdf(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Bayesian/figures/",method,"/",model,"/post_distrib",trait,"_GO_",method,"_",model,"_Taxus.pdf"));plot(pp_check(fit1));dev.off()

plot(conditional_effects(fit1), points = TRUE)
    
    pdf(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Bayesian/figures/",method,"/",model,"/lm_plot",trait,"_GO_",method,"_",model,"_Taxus.pdf"));plot(conditional_effects(fit1), points = TRUE);dev.off()
  }
}
# Print the correlation data frame
print(lm_bayesian_df)
```


```{r save lm bayesian model, include=FALSE}
#lm_Bayesian_df_GO_taxus<- lm_bayesian_df
#save(lm_inf_df_GO_taxus,file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Bayesian/table/lm_Bayesian_df_GO_taxus.Rdata")

load("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Bayesian/table/lm_Bayesian_df_GO_taxus.Rdata")
#save to table
pdf("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Bayesian/table/table_models.pdf", height=16, width=11.5)
grid.table(lm_inf_df_GO_taxus)
ttheme_default(base_size = 18)
dev.off()
```

We found similar results with the Bayesian approach. Overall, the best associations align with the findings from the Pearson correlation: both the RDA and GF models appear to be negatively associated with all the traits, with the strength of the association varying depending on the model and the trait. Furthermore, the composite index seems to be more strongly negatively associated with genomic offset compared to other traits.

        
## Non linear models
        
Due to the complexity of the relationship between the traits and the genomic offset predictions, we may not expect a linear relationship. To test this, we calculated non-linear models with quadratic parameters to capture any potential non-linear responses. We employed both inferential and Bayesian approaches for this analysis.

### Inferential model

We performed the model using the **nls** function from **stats** package: 
```{r non-linear frequentist model, echo=T, eval=F}
list_GO_2012_RDA <- c("GO_RDA_all_CG2012_Taxus","GO_RDA_random_CG2012_Taxus", "GO_RDA_random_same_AF_CG2012_Taxus", 
                      "GO_RDA_LC_CG2012_Taxus")#, "GO_RDA_CG_CG2012_Taxus"
list_GO_2021_RDA <- c("GO_RDA_all_CG2021_Taxus","GO_RDA_random_CG2021_Taxus", "GO_RDA_random_same_AF_CG2021_Taxus", 
                      "GO_RDA_LC_CG2021_Taxus")#, "GO_RDA_CG_CG2021_Taxus"

# List of Gradient Forest genomic offsets for 2012 and 2021
list_GO_2012_GF <- c("GO_standard_GF_all_2012","GO_standard_GF_random_2012", "GO_standard_GF_random_same_AF_2012", 
                     "GO_standard_GF_LC_2012")#, "GO_standard_GF_CG_2012"
list_GO_2021_GF <- c("GO_standard_GF_all_2021","GO_standard_GF_random_2021", "GO_standard_GF_random_same_AF_2021", 
                     "GO_standard_GF_LC_2021")#, "GO_standard_GF_CG_2021"

# Combine lists into one for easier processing
list_GO_2012 <- c(list_GO_2012_RDA, list_GO_2012_GF)
list_GO_2021 <- c(list_GO_2021_RDA, list_GO_2021_GF)

# Corresponding methods and models
list_method <- c(rep("RDA", 4), rep("GF", 4))  # 4 RDA models and 4 GF models

list_model <- c("all","random", "random_same_AF", "LC", "all","random", "random_same_AF", "LC")#, "CG", "CG"

# Traits considered
list_traits <- c("growth","growth_pheno","repro_pheno","leafthickness","composite_fitness")

# Create an empty dataframe to store correlations
nl_infer_df_taxus <- data.frame(
  Method = character(),
  Model = character(),
  Trait = character(),
  R2_model = numeric(),
  coef_b1 = numeric(),
  b1_Pvalue = numeric(),
  coef_b2 = numeric(),
  b2_Pvalue = numeric(),
  stringsAsFactors = FALSE
)

# Loop over each Genomic Offset for both RDA and GF models
for(i in 1:8) {  # Now we have 8 models (4 RDA + 4 GF)
  GO_pred_2012 <- get(list_GO_2012[i])
  GO_pred_2021 <- get(list_GO_2021[i])
  method <- list_method[i]
  model <- list_model[i]
  
  # Loop over each trait
 for(x in 1:length(list_traits)) {
    trait <- list_traits[x]
    
    # Use CG2021 data for Height, and CG2012 for other traits
    if (trait == "leafthickness" || trait== "composite_fitness" ) {
      df_trait <- get(paste0("df_", trait))
      
      # Merge data frames based on population for CG2021
      merged_df <- merge(df_trait, GO_pred_2021, by = "Population", suffixes = c(".trait", ".GO"))
    } else {
      df_trait <- get(paste0("df_", trait))
      
      # Merge data frames based on population for CG2012
      merged_df <- merge(df_trait, GO_pred_2012, by = "Population", suffixes = c(".trait", ".GO"))
    }

  y_data <- merged_df[,2]
  x_data <- merged_df[,3]
   
df_nl_model <- data.frame(y=y_data,x=x_data)

infer_model <- stats::nls(y ~ b0+b1*x+b2*(x^2),data=df_nl_model,
                          start = list(b0=0.1,b1 = 0.1, b2 = .1),control = nls.control(minFactor = 1e-10, maxiter = 500))

#save the model: 
assign(paste0("nl_infer_quadra_",method,"_",model,"_",trait),infer_model)

save(list = paste0("nl_infer_quadra_",method,"_",model,"_",trait),file=paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Non_linear/Inferential/models/",method,"/",model,"/nl_infer_quadra_",method,"_",model,"_",trait,".Rdata"))


# Extract the estimates for b1 and b2

coef_b1 <- data.frame(coef(infer_model)["b1"])
coef_b2 <- data.frame(coef(infer_model)["b2"])
R2_model <- modelr::rsquare(infer_model, df_nl_model)


#pvalues
model_summary <- summary(infer_model)
p_values_b1 <- coef(model_summary)["b1", "Pr(>|t|)"]
p_values_b2 <- coef(model_summary)["b2", "Pr(>|t|)"]


nl_infer_df_taxus <- rbind(nl_infer_df_taxus, data.frame(
      Method = method,
      Model = model,
      Trait = trait,
      R2_model = round(R2_model,3),
      coef_b1 = round(coef_b1[,1],3),
      b1_Pvalue = round(p_values_b1,3),
      coef_b2 = round(coef_b2[,1],3),
      b2_Pvalue = round(p_values_b2,3)
    ))

#confidence interval for the plot: 

ci <- confint(infer_model)
print(ci)

# Create new data for prediction over a range of x values
new_data <- data.frame(x = seq(min(df_nl_model$x), max(df_nl_model$x), length.out = 100))

# Get the predicted values for the new data
new_data$y_pred <- predict(infer_model, newdata = new_data)

# Bootstrapping to calculate confidence intervals for predictions
# Define the prediction function for the bootstrapping
predict_nls <- function(data, indices) {
        tryCatch({
          fit <- nls(y ~ b1*x + b2*(x^2), data = data[indices, ], start = list(b1 = 0.1, b2 = 0.1))
          predict(fit, newdata = new_data)
        }, error = function(e) {
          rep(NA, nrow(new_data))  # Return NA if fitting fails
        })
      }

boot_res <- boot::boot(
        data = df_nl_model,    
        statistic = predict_nls,  
        R = 1000
      )

# Filter out NA values before calculating quantiles
      valid_bootstrap <- boot_res$t[complete.cases(boot_res$t), ]
      
      # Calculate confidence intervals from bootstrapped predictions
      ci_low <- apply(valid_bootstrap, 2, quantile, probs = 0.025, na.rm = TRUE)
      ci_high <- apply(valid_bootstrap, 2, quantile, probs = 0.975, na.rm = TRUE)

# Add confidence intervals to the data
new_data$ci_low <- ci_low
new_data$ci_high <- ci_high

# Plotting the model with confidence intervals using ggplot2

plot <- ggplot(df_nl_model, aes(x = x, y = y)) +
  geom_point(color = "blue") +  # Observed data points
  geom_line(data = new_data, aes(x = x, y = y_pred), color = "red") +  # Fitted curve
  geom_ribbon(data = new_data, aes(y = x, ymin = ci_low, ymax = ci_high), alpha = 0.2, fill = "lightgrey") +  # Confidence interval band
  labs(title = "Non-Linear Model with Confidence Intervals",
       x = "x",
       y = "y") +
  theme_minimal()
plot
    
    pdf(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Non_linear/Inferential/figures/",method,"/",model,"/plot_nl_inf_quadra_",method,"_",model,"_",trait,"_Taxus.pdf"));plot(plot);dev.off()
  }
}
# Print the correlation data frame
print(lm_inf_df)
```


```{r non-linear relationship example rmarkdown, echo=FALSE}
for(i in 1:1) {  # Now we have 8 models (4 RDA + 4 GF)
  GO_pred_2012 <- get(list_GO_2012[i])
  GO_pred_2021 <- get(list_GO_2021[i])
  method <- list_method[i]
  model <- list_model[i]
  
  list_traits <- c("composite_fitness")
  # Loop over each trait
 for(x in 1:length(list_traits)) {
    trait <- list_traits[x]
    
    # Use CG2021 data for Height, and CG2012 for other traits
    if (trait == "leafthickness" || trait== "composite_fitness" ) {
      df_trait <- get(paste0("df_", trait))
      
      # Merge data frames based on population for CG2021
      merged_df <- merge(df_trait, GO_pred_2021, by = "Population", suffixes = c(".trait", ".GO"))
    } else {
      df_trait <- get(paste0("df_", trait))
      
      # Merge data frames based on population for CG2012
      merged_df <- merge(df_trait, GO_pred_2012, by = "Population", suffixes = c(".trait", ".GO"))
    }

  y_data <- merged_df[,2]
  x_data <- merged_df[,3]
   
df_nl_model <- data.frame(y=y_data,x=x_data)

infer_model <- stats::nls(y ~ b0+b1*x+b2*(x^2),data=df_nl_model,
                          start = list(b0=0.1,b1 = 0.1, b2 = .1),control = nls.control(minFactor = 1e-10, maxiter = 500))

# Extract the estimates for b1 and b2

coef_b1 <- data.frame(coef(infer_model)["b1"])
coef_b2 <- data.frame(coef(infer_model)["b2"])
R2_model <- modelr::rsquare(infer_model, df_nl_model)


#pvalues
model_summary <- summary(infer_model)
p_values_b1 <- coef(model_summary)["b1", "Pr(>|t|)"]
p_values_b2 <- coef(model_summary)["b2", "Pr(>|t|)"]

#confidence interval for the plot: 

ci <- confint(infer_model)
print(ci)

# Create new data for prediction over a range of x values
new_data <- data.frame(x = seq(min(df_nl_model$x), max(df_nl_model$x), length.out = 100))

# Get the predicted values for the new data
new_data$y_pred <- predict(infer_model, newdata = new_data)

# Bootstrapping to calculate confidence intervals for predictions
# Define the prediction function for the bootstrapping
predict_nls <- function(data, indices) {
        tryCatch({
          fit <- nls(y ~ b1*x + b2*(x^2), data = data[indices, ], start = list(b1 = 0.1, b2 = 0.1))
          predict(fit, newdata = new_data)
        }, error = function(e) {
          rep(NA, nrow(new_data))  # Return NA if fitting fails
        })
      }

boot_res <- boot::boot(
        data = df_nl_model,    
        statistic = predict_nls,  
        R = 1000
      )

# Filter out NA values before calculating quantiles
      valid_bootstrap <- boot_res$t[complete.cases(boot_res$t), ]
      
      # Calculate confidence intervals from bootstrapped predictions
      ci_low <- apply(valid_bootstrap, 2, quantile, probs = 0.025, na.rm = TRUE)
      ci_high <- apply(valid_bootstrap, 2, quantile, probs = 0.975, na.rm = TRUE)

# Add confidence intervals to the data
      
#new_data$ci_low <- ci_low
#new_data$ci_high <- ci_high

# Plotting the model with confidence intervals using ggplot2

plot <- ggplot(df_nl_model, aes(x = x, y = y)) +
  geom_point(color = "blue") +  # Observed data points
  geom_line(data = new_data, aes(x = x, y = y_pred), color = "red") +  # Fitted curve
  geom_ribbon(data = new_data, aes(y = x, ymin = ci_low, ymax = ci_high), alpha = 0.2, fill = "lightgrey") +  # Confidence interval band
  labs(title = "Non-Linear Model with Confidence Intervals",
       x = "x",
       y = "y") +
  theme_minimal()
plot
    # Generate the ggplot
    p <- ggplot(df_nl_model, aes(x = x, y = y)) +
      geom_point(color = "#FF1493") +
      geom_line(data = new_data, aes(x = x, y = y_pred), color = "blue") +  # Fitted curve
      geom_ribbon(data = new_data, aes(y = x, ymin = ci_low, ymax = ci_high), alpha = 0.2, fill = "lightgrey") +  # Confidence interval band
  labs(title = "Non-Linear Model composite fitness ~ Genomic offset GF random same AF ",
       x = "Genomic offset GF random same AF",
       y = "composite fitness index") +
  theme_minimal()+
      theme_bw(base_size = 11) +
  theme(legend.position="right", panel.grid = element_line(), strip.text = element_text(size=11),plot.title = element_text(hjust = 0.5,color = "Black",face="italic"))
    
    plot(p)
 }
}
#pdf("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Inferential/plot_model/Retained_model/totalstrob_RDA_random_AF_linear_plot.pdf");print(p);dev.off()
```

```{r save nl inf model standard GO, include=FALSE}
#df_nl_quadra_inf_taxus<- nl_infer_df_taxus
#save(df_nl_quadra_inf_taxus,file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Non_linear/Inferential/table/df_nl_quadra_inf_taxus.Rdata")

load("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Non_linear/Inferential/table/df_nl_quadra_inf_taxus.Rdata")

#save to table
pdf("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Non_linear/Inferential/table/table_models.pdf", height=16, width=11.5)
grid.table(df_nl_quadra_inf_taxus)
ttheme_default(base_size = 12)
dev.off()
```

### Bayesian model
```{r non-linear bayesian model, echo=T,eval=F}
list_GO_2012_RDA <- c("GO_RDA_all_CG2012_Taxus","GO_RDA_random_CG2012_Taxus", "GO_RDA_random_same_AF_CG2012_Taxus", 
                      "GO_RDA_LC_CG2012_Taxus")#, "GO_RDA_CG_CG2012_Taxus"
list_GO_2021_RDA <- c("GO_RDA_all_CG2021_Taxus","GO_RDA_random_CG2021_Taxus", "GO_RDA_random_same_AF_CG2021_Taxus", 
                      "GO_RDA_LC_CG2021_Taxus")#, "GO_RDA_CG_CG2021_Taxus"

# List of Gradient Forest genomic offsets for 2012 and 2021
list_GO_2012_GF <- c("GO_standard_GF_all_2012","GO_standard_GF_random_2012", "GO_standard_GF_random_same_AF_2012", 
                     "GO_standard_GF_LC_2012")#, "GO_standard_GF_CG_2012"
list_GO_2021_GF <- c("GO_standard_GF_all_2021","GO_standard_GF_random_2021", "GO_standard_GF_random_same_AF_2021", 
                     "GO_standard_GF_LC_2021")#, "GO_standard_GF_CG_2021"

# Combine lists into one for easier processing
list_GO_2012 <- c(list_GO_2012_RDA, list_GO_2012_GF)
list_GO_2021 <- c(list_GO_2021_RDA, list_GO_2021_GF)

# Corresponding methods and models
list_method <- c(rep("RDA", 4), rep("GF", 4))  # 4 RDA models and 4 GF models

list_model <- c("all","random", "random_same_AF", "LC", "all","random", "random_same_AF", "LC")#, "CG", "CG"

# Traits considered
list_traits <- c("growth","growth_pheno","repro_pheno","leafthickness","composite_fitness")

# Create an empty dataframe to store correlations

df_nl_bayesian_taxus <- data.frame(
  Method = character(),
  Model = character(),    # To store RDA name
  Trait = character(),  # To store trait name
  R2_model =numeric(),
  R2_ci_lower=numeric(),
  R2_ci_upper=numeric(),
  b1 = numeric(),       # To store b1 coefficient
  b2 = numeric(),       # To store b2 coefficient
  CI_b1 = numeric(),
  CI_b2 = numeric(),
  stringsAsFactors = FALSE
)

# Loop over each Genomic Offset for both RDA and GF models
for(i in 1:8) {  # Now we have 8 models (4 RDA + 4 GF)
  GO_pred_2012 <- get(list_GO_2012[i])
  GO_pred_2021 <- get(list_GO_2021[i])
  method <- list_method[i]
  model <- list_model[i]
  
  # Loop over each trait
 for(x in 1:length(list_traits)) {
    trait <- list_traits[x]
    
    # Use CG2021 data for Height, and CG2012 for other traits
    if (trait == "leafthickness" || trait== "composite_fitness" ) {
      df_trait <- get(paste0("df_", trait))
      
      # Merge data frames based on population for CG2021
      merged_df <- merge(df_trait, GO_pred_2021, by = "Population", suffixes = c(".trait", ".GO"))
    } else {
      df_trait <- get(paste0("df_", trait))
      
      # Merge data frames based on population for CG2012
      merged_df <- merge(df_trait, GO_pred_2012, by = "Population", suffixes = c(".trait", ".GO"))
    }

  y_data <- merged_df[,2]
  x_data <- merged_df[,3]
   
df_nl_model <- data.frame(y=y_data,x=x_data)
prior1 <- prior(normal(0, 10), nlpar = "b0")+prior(normal(0, 10), nlpar = "b1")+ prior(normal(0, 10), nlpar = "b2")

fit1 <- brm(bf(y ~ b0+b1*x+b2*(x^2), b0~1,b1 ~ 1,b2 ~ 1, nl = TRUE),
            data = df_nl_model, prior = prior1, iter=1500000,warmup = 50000,thin=500, cores = 16)

#save the model: 
#assign(paste0("nl_bayesian_quadra_",method,"_",model,"_",trait),fit1)

#save(list = paste0("nl_bayesian_quadra_",method,"_",model,"_",trait),file=paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Non_linear/Bayesian/models/",method,"/nl_bayesian_quadra_",method,"_",model,"_",trait,".Rdata"))

load(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Non_linear/Bayesian/models/",method,"/nl_bayesian_quadra_",method,"_",model,"_",trait,".Rdata"))

fit1 <- get(paste0("nl_bayesian_quadra_",method,"_",model,"_",trait))
#coeff

fit1_summary <- summary(fit1)

# Extract the estimates for b1 and b2
coef_b1 <- fit1_summary$fixed["b1_Intercept", "Estimate"]
coef_b2 <- fit1_summary$fixed["b2_Intercept", "Estimate"]

b1_ci_lower <- fit1_summary$fixed["b1_Intercept", "l-95% CI"]
b1_ci_upper <- fit1_summary$fixed["b1_Intercept", "u-95% CI"]
b2_ci_lower <- fit1_summary$fixed["b2_Intercept", "l-95% CI"]
b2_ci_upper <- fit1_summary$fixed["b2_Intercept", "u-95% CI"]

R2_model <- bayes_R2(fit1)["R2","Estimate"]

R2_ci_lower <- bayes_R2(fit1)["R2", "Q2.5"]
R2_ci_upper <- bayes_R2(fit1)["R2", "Q97.5"]



df_nl_bayesian_taxus <- rbind(df_nl_bayesian_taxus, data.frame(
      Method = method,
      Model = model,
      Trait = trait,
      R2_model= round(R2_model,3),
      R2_ci_lower=round(R2_ci_lower,3),
      R2_ci_upper=round(R2_ci_upper,3),
      b1 = round(coef_b1,3),
      b2 = round(coef_b2,3),
      CI_b1=paste0(round(b1_ci_lower,3),"_",round(b1_ci_upper,3)),
      CI_b2=paste0(round(b2_ci_lower,3),"_",round(b2_ci_upper,3))
    ))

#MCMC chains
plot(fit1)

#pdf(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Non_linear/Bayesian/figures/",method,"/",model,"/MCMC_nl_quadra_bayesian",trait,"_GO_",method,model,"_Taxus.pdf"));plot(fit1);dev.off()

#posterior distrib
pp_check(fit1)

pdf(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Non_linear/Bayesian/figures/",method,"/",model,"/post_distrib_nl_quadra_bayesian",trait,"_GO_",method,model,"_Taxus.pdf"));plot(pp_check(fit1));dev.off()


plot(conditional_effects(fit1), points = TRUE)
    
pdf(paste0("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Non_linear/Bayesian/figures/",method,"/",model,"/plot_nl_quadra_",trait,"_GO_",method,model,"_Taxus.pdf"));plot(conditional_effects(fit1), points = TRUE);dev.off()
  }
}

# Print the correlation data frame
print(df_nl_bayesian_taxus)
```


```{r save df result nl bayesian standard GO, include=FALSE}

#df_nl_bayesian_model_standard_GO_taxus<- df_nl_bayesian_taxus
#save(df_nl_bayesian_model_standard_GO_taxus,file="C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Non_linear/Bayesian/table/df_nl_bayesian_model_standard_GO_taxus.Rdata")

load("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Non_linear/Bayesian/table/df_nl_bayesian_model_standard_GO_taxus.Rdata")

#save to table
pdf("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Non_linear/Bayesian/table/table_models.pdf", height=16, width=11.5)
grid.table(df_nl_bayesian_model_standard_GO_taxus)
ttheme_default(base_size = 7)
dev.off()
```


Overall, we can observe that the relationships appear quite linear, as adding a quadratic parameter did not significantly change the R2 of the models. This suggests that the relationships may indeed be linear; therefore, incorporating a quadratic coefficient did not substantially impact the results as long as a linear coefficient remained in the model. Moreover, Bayesian and inferential models converged for those models with the highest negative associations.
**In conclusion, our dataset reveals a negative association between fitness proxies and genomic offset predictions. This suggests that genomic offset may provide valuable insights into the maladaptation of English Yew populations to climate conditions.**

# Draft

## other phenotypic traits

```{r other phenotypic traits, eval=FALSE, include=FALSE}
#Basal area
#We remove the ACEBDO, LASENIA pop
df_growth_bis_final <- df_growth_bis %>%
  filter(!(Population %in% c("LASENIA")))

##LMA
df_LMA_final <- df_LMA %>%
  filter(!(Population %in% c("ACEBEDO", "ALCOY", "LASENIA","	QUESADA")))

##D13C
df_StomDens_final <- df_StomDens %>%
  filter(!(Population %in% c("ACEBEDO", "ALCOY", "LASENIA","	QUESADA")))

##D13C
df_D13C_final <- df_D13C %>%
 filter(!(Population %in% c("ACEBEDO", "ALCOY", "LASENIA","	QUESADA")))
```


## output linear relationship poster

```{r linear relationship poster, eval=FALSE, include=FALSE}
merged_df <- merge(df_strobtot, GO_RDA_random_same_AF_CG2012_Taxus, by = "Population", suffixes = c(".trait", ".GO"))

lm_model <- lm(merged_df[, 2] ~ merged_df[, 3])

summary(lm_model)
coeff <- data.frame(coef(lm_model))[2,1]
    result_model <- data.frame(anova(lm_model))
    r_squared <- summary(lm_model)$r.squared
    pvalues <- result_model[1,5]
    

plot_data <- data.frame(
      GO_values = merged_df[, 3],
      Blup = merged_df[, 2],
      fitted = lm_model$fitted.values
    )
    
    # Generate the ggplot
    p <- ggplot(plot_data, aes(x = GO_values, y = Blup)) +
      geom_point(color = "#FF1493") +
      geom_smooth(method = "lm", se = TRUE, level = 0.95, color = "lightgrey", alpha = 0.2) +  # Add confidence interval bands
      geom_line(aes(y = fitted), color = "blue") +
      labs(title = "Total number strobili = 0.01 - 0.12xGO + E",
           x = "Genomic offset ",
           y = "Total number strobili") +
      theme_minimal()+
      theme_bw(base_size = 11) +
  theme(legend.position="right", panel.grid = element_line(), strip.text = element_text(size=11),plot.title = element_text(hjust = 0.5,color = "Black",face="italic"))
    
    plot(p)
    
pdf("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/Genomic_offset/Validation/GO/all_models_period/Linear_models/Inferential/plot_model/Retained_model/totalstrob_RDA_random_AF_linear_plot.pdf");print(p);dev.off()
```


```{r coeff poster, eval=FALSE, include=FALSE}

 merged_df <- merge(df_openstrob_final, GO_standard_GF_random_2012, by = "Population", suffixes = c(".trait", ".GO"))
model <- lm(merged_df[,2]~merged_df[,3])

summary(model)

fitted = model$fitted.values
    
    # Generate the ggplot
    p <- ggplot(merged_df, aes(x = merged_df[,3], y = merged_df[,2])) +
      geom_point(color = "blue") +
      geom_smooth(method = "lm", se = TRUE, level = 0.95, color = "lightgrey", alpha = 0.2) +  # Add confidence interval bands
      geom_line(aes(y = fitted), color = "red") +
      labs(title = paste("Linear Model:", trait, "~", method, model),
           x = paste0("Genomic offset ",method," ",model),
           y = paste0("proxy fitness ",trait)) +
      theme_minimal()

    p
    
 df_nl_model<- data.frame(y=merged_df$Blups,x=merged_df$genomic_offset_GF_random)
    
infer_model <- stats::nls(y ~ b0+b1*x+b2*(x^2),data=df_nl_model,
                          start = list(b0=0.1,b1 = 0.1, b2 = .1),control = nls.control(minFactor = 1e-10, maxiter = 500))

coef_b1 <- data.frame(coef(infer_model)["b1"])
coef_b2 <- data.frame(coef(infer_model)["b2"])
R2_model <- modelr::rsquare(infer_model, df_nl_model)


#pvalues
model_summary <- summary(infer_model)
p_values_b1 <- coef(model_summary)["b1", "Pr(>|t|)"]
p_values_b2 <- coef(model_summary)["b2", "Pr(>|t|)"]
```

## non-parametric tests 

```{r non parametric test, eval=FALSE, include=FALSE}

spearman_test <- cor.test(merged_df[,3], merged_df[,2], method = "spearman")


kendall_test <- cor.test(merged_df[,3], merged_df[,2], method = "kendall")
```


