---
title: "Variance_partitioning"
author: "Thomas Francisco"
date: "2024-04-30"
output:
  html_document:
    number_sections: true #titles
    toc: true #table of content
    toc_float: true # enable the toc to be on the side of the text, always visible
    collapsed: True #control if the toc label will only display top level titles
    toc_depth: 3
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(cache = FALSE)

library(vegan)
library(dplyr)
library(adespatial)
library(geodist)
library(writexl)
library(ggplot2)
```


# Introduction 

The goal of this script is to perform variance partitioning to disentangle the variables that explain genetic variation. The second goal is to determine whether climatic variables account for some of the genomic variation, suggesting the presence of candidate loci for local adaptation to climatic variables across populations. Indeed, one of the assumptions of GEA models is that environmental variables explain part of the genetic variation, and we aim to test this.  
Additionally, variance partitioning helps visualize the extent of covariation between population structure, IBD (Isolation by Distance), and the environment. Since there is no single best proxy for ancestry or IBD, we employed several proxies for population structure and geographic distance.  
To carry out this variance partitioning, we used RDA and pRDA analyses, which are explained in the script *Redundancy_analyses_candidate_detection*.  

# Data

## DBMEMs analysis

To estimate geographical differences, we can use the geographical coordinates, but we can also use a more complex index such as a matrix of geographical correlation along the study area between pops, such as the Distance Based Eigenvector Maps (dbMEM). Indeed, dbMEM will improve the patterns of spatial autocorrelation by finding more autocorrelation patterns and may be able to disentangle large and fine autocorrelation scales.

dbMEM can be used in two ways:  
- estimate the autocorrelation of a variable (e.g. temperature) based on the geographical distances between sampled points  
- the autocorrelation of the sampled points directly, without using the explanatory variables.
                                         
 We calculated the autocorrelation using DBMEM. 
 
```{r initial data}
meta_data_pop <- read.csv("C:/Users/tfrancisco/Documents/Thèse/Data/Espèces/Taxus_baccata/Populations/taxus_sample_29pop.csv",h=T,sep=";",dec=",")

#data frame with only the coordinates
coord_data <- meta_data_pop[,c(4:5)] %>% 
  apply(2,as.numeric) %>% 
  data.frame()
row.names(coord_data) <-  meta_data_pop$Population
```

## Distance matrix
Our input variables of the distance matrix are coordinates, we can calculate 2 types of distance matrices: 
euclidean: distance between two points in 2 dimensions
geodesic: distance between two points in 3 dimensions -> take into account the curvature of the earth
We've used the geodesic distance because it's more accurate. 

```{r geodesic distance}
distance_matrice <- geodist(coord_data, measure= "geodesic") 
dist_matrix <- as.dist(distance_matrice)
```

## dbMEM model
We then ran DBMEM on the distance matrix. 
We kept only the positive correlation to be used in variance partitioning, and the threshold for truncation of the geodesic distance is the length of the longest edge of the minimum spanning tree.
We did not perform a selection of eigenvectors (dbMEM axes) based on p-values, we kept all dbMEM axes with positive correlation.
```{r dbMEM function}
dbmem_results <- dbmem(dist_matrix, MEM.autocor = "positive")

# Plot spatial map for each eigenvector
plot(dbmem_results[,3])

meta_data_pop$Latitude <- as.numeric(meta_data_pop$Latitude)
meta_data_pop$Longitude <- as.numeric(meta_data_pop$Longitude)


plots <- lapply(1:ncol(dbmem_results), function(i) {
  ggplot() +
    geom_point(data = meta_data_pop, aes(x = Longitude, y = Latitude, color = dbmem_results[, i])) +
    scale_color_gradientn(colours = c("black", "grey", "red"), values = scales::rescale(c(-2, 0, 3))) +
    labs(title = paste("Eigenvector", i),color = "Eigenvalues") +
    theme_minimal()
  
})
pdf("C:/Users/tfrancisco/Documents/Thesis/Results/species/taxus/DBMEM/DBMEMaxes.pdf");print(plots);dev.off()

# Print the plots
plots
```

Interpretation:  
DBMEM1: longitudinal gradient  
DBMEM2: extreme latitudes in the same group and intermediate latitudes in a different group  
DBMEM3 and 4: less clear  

Save DBMEMs in a meta-file to be used in variance partitioning and GEA if variables are relevant to explain genetic variation. 
```{r add DBMEMs to metadata}
#final dataset
meta_data_dbmem_29pop_adapcon_gentree <- data.frame(meta_data_pop,dbmem_results)

meta_data_dbmem_29pop_adapcon_gentree_scale <- meta_data_dbmem_29pop_adapcon_gentree %>% 
    mutate(across(where(is.numeric), scale))

#exportation
#write_xlsx(meta_data_dbmem_29pop_adapcon_gentree_scale,"C:/Users/tfrancisco/Documents/Thèse/Data/Espèces/Taxus_baccata/GEA/dbMEMs/meta_data_dbmem_29pop_adapcon_gentree_scale.xlsx")
#save(meta_data_dbmem_29pop_adapcon_gentree_scale,file="C:/Users/tfrancisco/Documents/Thèse/Data/Espèces/Taxus_baccata/GEA/dbMEMs/meta_data_dbmem_29pop_adapcon_gentree_scale.Rdata",force=T)
```
 
## Datasets 

Secondly, we load the other datasets. Since the goal is to disentangle the effects of ancestry, climate, and IBD (Isolation by Distance) on genetic variation, we used proxies for both ancestry and IBD. These proxies allow us to account for population structure and geographic factors when analyzing the genetic data.
```{r intial data}
#climatic data
##0.6 corr
Past_climatic_new_var <- read.csv("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/Climatic_data/new_selection/Past_new_6_Climatic_data_scale_df.csv",sep=";",dec=",") 

#Ancestry proxies
##PCs axis
structure_PCs <- read.csv("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/GEA/Structure_proxy/PCa_score_T_adapcon_gentree_scale.csv",sep=";",dec=",")

##STRUCTURE Groups
load("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/GEA/Structure_proxy/Cluster_score_STRUCTURE_T_Adapcon_gentree.Rdata")
Structure_cluster <- Cluster_score_STRUCTURE_T_Adapcon_gentree

#IBD proxy
#dbMEMs
IBD_dbMEMs <- read.csv("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/GEA/dbMEMs/meta_data_dbmem_29pop_adapcon_gentree_scale.csv",sep=";",dec=",")
#genomic data
load("C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/genetic_new/data_allelic_frequencies_29pop_adapcon_gentree_475_8616.Rdata")
genomic_matrix <- data_allelic_frequencies_29pop_adapcon_gentree_475_8616
```

```{r merge all informations into 1 table}
#new_var
Data_RDA <- merge(Past_climatic_new_var,structure_PCs[,c(1:3)],"Population") %>% merge(Structure_cluster,"Population") %>% merge(IBD_dbMEMs[,c(2,4,5,7:10)],"Population") 

#scale coord
scale_coord <- Data_RDA[,c(14,15)] %>%
  apply(2,as.numeric) %>% scale()

Climatic_data_RDA_pRDA <- data.frame(Data_RDA[,-c(14,15)],scale_coord)

#Save for RDA, pRDA
save(Climatic_data_RDA_pRDA,file="C:/Users/tfrancisco/Documents/Thesis/Data/Species/Taxus_baccata/GEA_new_var/variance_partitioning/Climatic_data_RDA_pRDA.Rdata")
```

# Models

We performed several models with different proxies for the population structure and the geography

## Model: coords as IBD proxy and PCs as ancestry proxy
```{r variance partitionning with coord as IBD proxy and PCs as ancestry proxy}
#total model
M_tot_coord <- rda(formula=genomic_matrix~Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality+PC1+PC2+Longitude+Latitude,data = Climatic_data_RDA_pRDA,scale=T)
#RsquareAdj(M_tot_coord)
#anova.cca(M_tot_coord)

##partial_ancestry_model
M_ancestry_coord <- rda(formula = genomic_matrix ~ PC1+PC2 +Condition(Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality +Longitude+Latitude), data = Climatic_data_RDA_pRDA,scale=T)
#RsquareAdj(M_ancestry_coord)
#anova.cca(M_ancestry_coord)

#partial_geo_model
M_geo_coord <- rda(formula = genomic_matrix ~ Longitude+Latitude+ Condition(Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality+PC1+PC2), data = Climatic_data_RDA_pRDA,scale=T)
#RsquareAdj(M_geo_coord)
#anova.cca(M_geo_coord)

#partial_env_RDA
M_IBE_coord <- rda(formula = genomic_matrix ~  Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality +Condition(PC1+PC2+Longitude+Latitude) , data = Climatic_data_RDA_pRDA,scale=T)
RsquareAdj(M_IBE_coord)
##summary(M_IBE_coord)
#anova.cca(M_IBE_coord)
```
Interpretation:   
R^2 tot     : 0.52                     Adj R^2 tot     : 0.25    
R^2 ancestry: 0.13                     Adj R^2 ancestry: 0.10  
R^2 IBD     : 0.06                     Adj R^2 IDB     : 0.01  
R^2 climate : 0.18                     Adj R^2 climate : 0.02  

## Model: dbMEMs as IBD proxy and Structure groups as ancestry proxy
```{r variance partitionning with mscore_structure as Structure proxy and dbMEMs as IBD proxy}
#total model
M_tot_db <- rda(formula=genomic_matrix~Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality+Group1+Group2+Group3+MEM1+MEM2+MEM3+MEM4,data = Climatic_data_RDA_pRDA,scale=T)
#RsquareAdj(M_tot_db)
#anova.cca(M_tot_db)

##partial_ancestry_model
M_ancestry_db <- rda(formula = genomic_matrix ~ Group1+Group2+Group3 +Condition(Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality +MEM1+MEM2+MEM3+MEM4), data = Climatic_data_RDA_pRDA,scale=T)
#RsquareAdj(M_ancestry_db)
#anova.cca(M_ancestry_db)

#partial_geo_model
M_geo_db <- rda(formula = genomic_matrix ~ MEM1+MEM2+MEM3+MEM4+ Condition(Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality+Group1+Group2+Group3), data = Climatic_data_RDA_pRDA,scale=T)
#RsquareAdj(M_geo_db)
#anova.cca(M_geo_db)

#partial_env_RDA
M_IBE_db <- rda(formula = genomic_matrix ~  Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality +Condition(Group1+Group2+Group3+ MEM1+MEM2+MEM3+MEM4) , data = Climatic_data_RDA_pRDA,scale=T)
#RsquareAdj(M_IBE_db)
#summary(M_IBE_db)
#anova.cca(M_IBE_db)
```

Interpretation:   
R^2 tot     : 0.60                     Adj R^2 tot     : 0.26    
R^2 ancestry: 0.12                     Adj R^2 ancestry: 0.07  
R^2 IBD     : 0.13                     Adj R^2 IDB     : 0.04  
R^2 climate : 0.19                     Adj R^2 climate : 0.04  

## Model: coords as IBD proxy and Structure groups as ancestry proxy
```{r variance partitionning with coord as IBD proxy and Structure as ancestry proxy}
#total model
M_tot_coord <- rda(formula=genomic_matrix~Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality+Group1+Group2+Group3+Longitude+Latitude,data = Climatic_data_RDA_pRDA,scale=T)
#RsquareAdj(M_tot_coord)
#anova.cca(M_tot_coord)

##partial_ancestry_model
M_ancestry_coord <- rda(formula = genomic_matrix ~ Group1+Group2+Group3 +Condition(Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality +Longitude+Latitude), data = Climatic_data_RDA_pRDA,scale=T)
#RsquareAdj(M_ancestry_coord)
#anova.cca(M_ancestry_coord)

#partial_geo_model
M_geo_coord <- rda(formula = genomic_matrix ~ Longitude+Latitude+ Condition(Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality+Group1+Group2+Group3), data = Climatic_data_RDA_pRDA,scale=T)
#RsquareAdj(M_geo_coord)
#anova.cca(M_geo_coord)

#partial_env_RDA
M_IBE_coord <- rda(formula = genomic_matrix ~  Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality +Condition(Group1+Group2+Group3+Longitude+Latitude) , data = Climatic_data_RDA_pRDA,scale=T)
#RsquareAdj(M_IBE_coord)
#summary(M_IBE_coord)
#anova.cca(M_IBE_coord)
```
Interpretation:   
R^2 tot     : 0.54                     Adj R^2 tot     : 0.24    
R^2 ancestry: 0.15                     Adj R^2 ancestry: 0.09  
R^2 IBD     : 0.06                     Adj R^2 IDB     : 0.01  
R^2 climate : 0.18                     Adj R^2 climate : 0.03  


## Model: dbMEMs as IBD proxy and PCs as ancestry proxy
```{r variance partitionning with dbMEMs as IBD proxy and PCs as ancestry proxy}
#total model
M_tot_db <- rda(formula=genomic_matrix~Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality+PC1+PC2+MEM1+MEM2+MEM3+MEM4,data = Climatic_data_RDA_pRDA,scale=T)
RsquareAdj(M_tot_db)
anova.cca(M_tot_db)

##partial_ancestry_model
M_ancestry_db <- rda(formula = genomic_matrix ~ PC1+PC2 +Condition(Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality +MEM1+MEM2+MEM3+MEM4), data = Climatic_data_RDA_pRDA,scale=T)
RsquareAdj(M_ancestry_db)
anova.cca(M_ancestry_db)

#partial_geo_model
M_geo_db <- rda(formula = genomic_matrix ~ MEM1+MEM2+MEM3+MEM4+ Condition(Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality+PC1+PC2), data = Climatic_data_RDA_pRDA,scale=T)
RsquareAdj(M_geo_db)
anova.cca(M_geo_db)

#partial_env_RDA
M_IBE_db <- rda(formula = genomic_matrix ~  Annual_Tc+Diurnal_range_Tc+Tc_Seasonality+Tc_driest_quarter+Annual_P+P_Seasonality +Condition(PC1+PC2+ MEM1+MEM2+MEM3+MEM4) , data = Climatic_data_RDA_pRDA,scale=T)
RsquareAdj(M_IBE_db)
#summary(M_IBE_db)
anova.cca(M_IBE_db)
```

Interpretation:   
R^2 tot     : 0.59                     Adj R^2 tot     : 0.28    
R^2 ancestry: 0.11                     Adj R^2 ancestry: 0.08  
R^2 IBD     : 0.13                     Adj R^2 IDB     : 0.04  
R^2 climate : 0.18                     Adj R^2 climate : 0.03  

**Conclusion: We retained the model using dbMEMs as the proxy for IBD and PCs as the proxy for ancestry. This model explains the most genetic variation, while maintaining a low number of variables, making it the most efficient and informative choice for our analysis.** 


